### 2015-03-12 rebuttal.txt
Initial import from SVN.

* Reviewer 1
** Anomaly detection
The paper describes an approach to fuse reports from IDSes; it does not do anomaly detection. MIFD fuses IDS reports from both anomaly- and signature-based IDSes. Perhaps confusion arises from System Z+'s use of the term "surprise" for its measure, which goes up as likelihood goes down.
** Code and data are not available
The data will be made available on the web. The code could be made available under limited license for research purposes, making reproduction and comparison feasible.
** Comparison
While this would be desirable, existing IDS fusion techniques do not share an input format or output definition. Most simply cluster and do not categorize hypotheses by posterior likelihood. However, it would be simple for other researchers to adapt their systems to our data sets: our synthetic reports would be simple to translate into other input formats.
* Reviewer 11
** Clarifications
*** Three sensors
You'd noted in the review that we "used three sensors for possible fusion." That's not exactly right --- although we do use a ratio of three sensors detecting each attack in most experiments, we vary that ratio in the first experiment.
*** Removing event hypotheses and sensor reports
We'd intended this remark as something of an aside, since (as the paper states) we are not actually doing this cleanup in the system under experimentation.  
The volume of data which IDSs encounter necessitates some sort of garbage collection. You are correct that the timing of such clean up must be carefully considered. In the full system, hypotheses are only purged after an interval since receipt of the last related IDS report. All reports and hypotheses are archived in a database.
It would probably be best to simply cut this minor aside from the paper, since garbage-collection plays no role in these experiments.
** Justification for choice of parameter settings
In all cases, the initial settings for the configuration parameters are both typical values for an IDS, and consistent with Z+'s assumptions. The experiments vary these values in ways which we anticipate would be worse for MIFD's performance, to examine our expected performance both under typical conditions, and under less typical conditions approaching worst-case.
** Relation to boosting
Boosting does provide an alternative method for fusion, but is not appropriate to IDS fusion. Boosting is based on learning, but IDS fusion problems do not admit use of ML: we do not get labeled data of attacks, nor do we get reference data showing systems operating without attacks. We will add a citation to papers about the failure of ML for IDS. These challenges (see [Sommer&Paxson,2010;Gates&Taylor,2006] on the challenges of ML in IDS, full cites available upon request) account for our adoption of a qualitative, non-adaptive framework. Also note that different IDSes do not share a single "ontology" of events; part of the function of MIFD's clustering is to align reports with different frames of reference.
** Other suggestions
Thank you for the many thoughtful suggestions for clarification.  We will edit the text to address all of these issues.
* Reviewer 4
** Novelty, technical quality, and significance
We respectfully disagree with this review's judgment about the novelty of our work.
It is true that System Z+ is well-studied in the literature, and /proposals/ to apply it to inference problems are common. However, we reviewed the 250+ papers citing the Z+ AIJ paper on Google Scholar (We will make the list of citations available upon request), and found no other paper describing an application of Z+ to a real-world problem. Accordingly, we assert that studying the application, and more to the point, the range of applicability, of System Z+ is novel.
Further, in order to apply an approximation technique such as Z+, one must have a clear understanding of when it will work, when it will fail, and how it will degrade from one to the other. Z+ is based on an idealization of uncertain reasoning, and it is necessary to study how it will work when its simplifying assumptions are violated. Our experiments studied exactly this --- beginning with assumptions in line with both the particular application of IDS and with Z+, and then diverging from them --- not "testing for the sake of testing."
There are numerous useful points of comparison in the literature: Consider how naive Bayesian classification has been studied: Anecdotal experience that useful results were computed even when violating independence assumptions led to more systematic studies of the integrity of Bayes classification as the independence failed.  Heckerman and Horvitz's study of why confidence factors work is in a similar vein.  Another point of comparison is big-O analysis: while it is of great utility, there are many cases where the constant factors do matter, and the idealization provides misleading results. Failure to investigate such cases would hamper computer science as a whole.
In this case System Z+ has been believed to be a useful approximation to real probabilities, and the novelty here is to examine the performance of Z+ as its assumptions are increasingly violated (with particular application to sensor fusion).
If we aren't willing to accept research that explores why a technique works or fails to work in practice, we doom applications work to be prescientific. Any single application can be made to work with a large enough investment of one-off labor: generalizing and validating requires follow-up.
Finally, we stress that our results are relevant to sensor fusion applications in general, not just IDS fusion.
We will improve the clarity of these introductory points in the final version of the paper.
** Quality of writing
Although it is difficult to address the non-specific claim of poor writing, our incorporation of the specific suggestions from the above reviews may address this reviewer's concerns.

* COMMENT Uses of System Z+
We have looked at the 250+ citations of G&P's AIJ paper on System Z+. Of these
citations, only the following show an application of the technique. And none are
to actual real-world applications. Accordingly, we can challenge the claim that
this is not novel.
+ Approximating MDP in [[http://pdf.aminer.org/000/387/692/faster_heuristic_search_algorithms_for_planning_with_uncertainty_and_full.pdf][Bonet and Geffner]].  Same technique used in
  [[http://juban.free.fr/hp/publications/challenge_pomdp.pdf][The Challenge of Solving POMDPs for Control, Monitoring and Repair of Complex Systems]].
+ Plan recognition, [[http://www.aaai.org/ocs/index.php/AAAI/AAAI10/paper/download/1821/2160][Ramirez&Geffner]]
+ (possibly) [[http://link.springer.com/chapter/10.1007/978-3-319-02621-3_11][Diagnosing Dependent Action Delays in Temporal Multiagent Plans]]

* COMMENT other notes:
+ From description of clustering:
  Note
  that clustering does not itself generate a single output; clustering
  is in fact the first step to generate all of the possible
  hypotheses. The separate step of assessment assigns qualitative
  probabilities to the hypotheses. Even then, assessment might not
  produce a /single/ output; it may in fact be more than one event
  which causes the observed reports, or we may hypothesize different
  events with equal likelihood.

  We'll describe the IDS fusion problem more generally.
+ Definition of MIFD in abstract.
+ Positioning of figures on the same page as referenced
+ Axis labels in Figures 3-7.
+ Readability of Figure 6.

** Reviewer 11's other suggestions
+ Defining benign event hypotheses vs. attack event hypotheses.
+ Defining complex event hypotheses. Briefly, these refer to the case
  where a hypothesized event is taken as evidence for another
  hypothesized event.  Since we do not use these in the current experiments,
  we may drop this remark from the final paper.
+ Better define the notation used in table of System Z+ rules. \omega
  is a possible world; \phi\in\omega is the same as \omega\models\phi.
+ Clustering. MIFD generates event hypotheses for incoming sensor
  reports based on its model of attacks and sensors.
  In general, clustering will generate multiple hypotheses for a single report,
  and multiple reports may support a single hypothesis.
+ Suggestions to the intro/related work balance in "Intrusion
  Detection Fusion" section, and comparison to boosting.

* COMMENT Possible note
They have, counting extremely generously, 2 application-type papers in UAI 2013,
out of a total of 73.  It's tempting to raise this, but I'm not sure that I
care, because I suspect I'd rather have this paper accepted at AAAI instead of
UAI.

# Local Variables:
# mode: org
# End:

### 2015-03-12 abstract.tex
Initial import from SVN.

%% WONTFIX - Robert, should we move the ``of Goldszmidt and Pearl'' up
%% to the first mention of \zplus ?
%% I don't think so: I think that introducing subordinate phrases like that
%% makes the topic sentence lose focus. We give them credit; that's all we need
%% to do.
\begin{abstract}
We experimentally analyze the accuracy of the \zplus qualitative
probability scheme of
Goldszmidt and Pearl when used for diagnosis and information fusion.
The \longids
(\ids) fusion system
Scyllarus, and its successor MIFD, use $Z+$ to assess the likelihood of various cyber attack events based on
reports from \idses.
$Z+$ provides an order of magnitude approximation of
conventional probability, similar to the order of magnitude approximation of
computational complexity provided by big-O analysis.
Scyllarus accurately identifies attacks and
substantially reduces the false positives that are the bane of intrusion detection.
% However,
% it is difficult to draw \emph{general} conclusions about an intrusion detection
% approach based on
% field testing, since variables are not systematically controlled.
In the work described here, we experimentally analyze the performance of MIFD
in order to provide general conclusions about its behavior, complementing the
results from field tests.
Our experiments show that the qualitative probability scheme degrades gracefully
in precision and recall
as its order of magnitude approximation is a less and less accurate
representation of true distributions.
The system also degrades gracefully as its input sensors become
less discriminating.
Finally, we show that qualitatively fusing multiple IDSes successfully
addresses base rate issues in intrusion detection.
The interest of these results is not limited to intrusion detection: the method
used in our systems is a general abductive scheme, based on qualitative Bayes
networks, so the results are applicable to other information fusion and
diagnostic applications.
To the best of our knowledge, ours is the only experimental
investigation of the accuracy of $Z+$ as an approximation of conventional probability.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 design.tex
Initial import from SVN.

\section{Experimental Design}
\label{sec:experimental-design}

Our experiments probe limits of accuracy of \zplus.
Our first two experiments consider what happens as our qualitative abstraction
is a more or less accurate reflection of reality.  It is relatively accurate when the
probabilities of events it treats as qualitatively distinct are far apart, and
less and less accurate as they come closer and closer together. \hide{This is similar
to the way big-$O$ analysis degrades as the constants get larger and larger.}
Our third experiment considers how MIFD degrades as its input sensors become
less and less precise.
Finally, we consider how well MIFD addresses base rate problems -- to what
extent, and under what circumstances fusing redundant sensors qualitatively will
increase accuracy.

In all of these experiments, we define sensor arrangements, and probabilities for attacks, confounding
benign events, and sensor false positives and false negatives. We sample from
them to find ground truth and sensor report sets, run MIFD on the sensor
reports, and evaluate its accuracy.
These experiments are necessarily artificial, since they assume
omniscience about parameters which are even in principle inaccessible.

\hide{
Our experiments are based on \emph{settings} comprised of
%% a set of  \emph{locations} (simulating hosts),
attack and benign event \emph{prototypes}, plus
\emph{sensors} which simulate \idses, each
sensing some number of attacks and possibly also (inadvertently)
benign events.  Since there is no opportunity to perform 
fusion if events are not sensed, in our simulated configurations each attack
event will be sensed by some number of sensors, greater than 1.  We vary this
sensor-to-attack ratio in our experiments.}
% In this first set of experiments, we do not simulate networks as
% such, so all of the locations are treated as identical: the same set
% of events can occur everywhere, and the same set of sensors are
% deployed at each location.

% DONE: Can you give the reader a sense of the distinction between
% configuration, setting, and run? The notion of "run" is relatively clear, but
% not the other two.  Something like "For example, a configuration might specify
% how a set of attacks is generated, and the set of sensors. A particular
% setting would involve a particular set of possible attacks, a particular set
% of sensors, and a particular set of sensing relations between attacks and
% sensors."
%
% FIXME - Robert, in the marked-up PDF you suggested discussing
% locations, but earlier we'd decided to prune that discussion, since
% we didn't end up using different locations.
% Right, but here how will the reader know that we are going to sample from the
% same event multiple times?  The fact is that the whole notion of "run" as
% distinct from "location" is broken, so we must make the best of it that we
% can.  Please have a whack at this.
For each experiment we create a \emph{configuration},
parameters specifying how a set of events and sensors (the
\emph{setting}) is generated.  A
setting comprises a set of possible attacks, a set
of sensors, and a set of sensing relations between attacks and
sensors.
In each \emph{run} of a setting we sample %, for each location,
% some number of 
from the attack and benign events.
% These are
% Bernoulli random variables; we discuss setting their probabilities
% below.
In each experiment we generated 1,000 settings with
100 runs per setting.
%
After sampling events, we sample from the sensors
%at the corresponding location,
according to their false positive and false negative probabilities.  
\hide{For each of
the $i$ events which the sensor might detect, we consider the probability
$\mathit{fn}$ of a false negative, so the probability of the sensor firing is $1
- \mathit{fn}^i$. If none of the sensed events occurred, the sensor will fire
according to its false positive probability.
}

\hide{
When measuring recall, we care only about trials where attacks
actually occur, but in some configurations the probability of an attack
is so small that the sample size required to meaningfully assess
recall with standard sampling would not be feasible.
Conversely, when measuring precision, false positives are of critical
importance, so samples without attacks are interesting.
Accordingly we sample separately for precision and recall, and
for recall we perform rejection sampling, rejecting any event samples
in which no attacks occur.}

Having generated the events and sensor reports for a run, we then use
the reports as
input to MIFD, and assess the resulting Bayes networks.  We extract
the set of attack event hypotheses that have been labeled as
\emph{likely} by MIFD.
We evaluate MIFD's performance in terms of \emph{precision} and \emph{recall},
comparing the set of events that MIFD considers likely with ground
truth generated for the run.
Recall is the percentage of actual attack events
which are labeled as likely. \hide{We compute recall as the fraction
\[\frac{\textrm{correct positives}}
       {\textrm{correct positives}+\textrm{false
           negatives}}\enspace.\] }
%
Precision is the percentage of attack events labeled as likely which 
actually occurred.\hide{
in fact part of the ground truth. We compute precision as the fraction
\[\frac{\textrm{correct positives}}
       {\textrm{correct positives}+\textrm{false positives}}\enspace.\]
%
We generally consider precision and recall aggregated over all runs
from a setting.}
Because of the high rate of reports and the low rate of events, precision and
recall must be in the high nineties, or the sensing system will not be usable.
We report the effect of our experiments on precision and recall in our
results section below.

\textbf{Configurations.}
\label{sec-5-1}
%% Explain how we generate settings, events, benign events, and sensor
%% reports.
% Generation of each experimental run of \mifd\ is governed by a number
% of experimental parameters.  
% Over any particular experiment, we will vary one or two parameters,
% and set others to values typical for an \ids.
In each of our experiments, we start with a configuration that reflects the
realities of IDS fusion, and whose parameter values are broadly consistent with
the simplifying assumptions of \zplus.
Configuration parameters may be either constants or parameterized random
variables.
We then vary some configuration parameter in ways which degrade
MIFD's performance. These experiments show us MIFD's performance under best-case conditions, and
how that performance degrades.
\begin{compactitem}
\item \numEventProtosDetected\ --- The total number of events (both
  benign and attacks) to be
  detected by each sensor, a measure of the sensor's specificity.
  Initially we use a random variable which returns 1 80\%
  of the time, and 2 otherwise.
\item \sensorToAttackRatio\ --- The number of sensors which should
  detect each attack event, initially 3.
\item \sensorOverlap\ --- The number of types of attack to be detected by each
  sensor, specified as a probability checked when deciding whether to
  associate an additional attack type with a sensor, initially 0.2.
\item \fpKappa\ --- The qualitative probability of
  false-positives for the sensors. We choose these to be 1
  in configurations where it takes 3 reports to rank a hypothesis likely, or 2
  where it takes only 2.
% 
%% Commenting out this bullet point --- not really relevant.
%
% \item $\kappa(\text{false-negative})$ --- MIFD does not currently reason about false negatives;
%  we take this probability to be whatever
%  corresponds to $\kappa = 2$ in the configuration.
%
\item \attackKappa\ and \benignKappa\ --- The qualitative
  probabilities of attack and benign events, initially
  2 and 1 respectively.

  %% This item is commented out in favor of the one below ---
  %% because we don't actually use \numBenignEvents in any experiment
  %% reported here.

  %% \item \numBenignEvents\ and \sensorToBenignRatio\ --- Exactly one
  %%   of these parameters should be given, and the other will be
  %%   derived. The former specifies a total number of benign events
  %%   which should be used in generated settings; the latter
  %%   specifies the number of sensors which should detect each benign
  %%   event.  We default to a constant value of 3 for
  %%   \sensorToBenignRatio.
\item \sensorToBenignRatio\ --- The number of sensors which should
  detect each benign event, initially 3.

\item \kappaTranslations\ --- Translations (into a constant or into a
  random variable) of qualitative probability values to real values in
  $[0,1]$. Initially we translate to constants, $\kappa(0)=0.5$,
  $\kappa(1)=0.01$, $\kappa(2)=0.001$.

% \item \numLocations --- Defaults to 1.

\item \numAttacks\ --- % Number of attack prototypes; 
  The number of attacks to take place, %at each location
  initially 4.

\end{compactitem}
%
% The values of \sensorToAttackRatio\ and the various \emph{kappa}s are
% interrelated. A sensible scenario should respect the orderings
% \begin{align}
% \end{align}
% or in other words,
% \begin{compactitem}
% \item A single sensor false positive is less surprising than an
%   attack event.
% \item A benign event is also less surprising than an attack event.
% \item An attack event is less surprising than false positives from all
%   of the sensors which detect that event.
% \end{compactitem}
%
%% (run-config ...)
%%  > (make-configuration-settings configuration)
%%  > (make-setting :configuration configuration)
%%  > (make-instance 'setting :configuration configuration ...)
%%    :after (initialize-instance ((obj setting)) ...)
%%             > (make-attack-events num-attacks attack-kappa)
%%                 ;; just the attack protos
%%    (setf (slot-value setting 'sensors) (make-sensors setting))
%%                 ;; sensor and benign protos
%%    (stratus-simulator:init-random-mifd-world setting num-locations
%%    ...)
%
%
\textbf{Settings.}
When creating a setting, we first consult
the \numAttacks\ parameter to determine the number of attack prototypes.
Each prototype is associated with an
occurrence probability given by (or sampled for each prototype from)
\attackKappa. The number of sensors is not fixed, but instead depends
on several configuration parameters. For each attack, we keep track of
the number of sensors which we must assign to detect it; this value is
initially set from \sensorToAttackRatio.  We create new sensors as
long as any attack requires assignment to an additional sensor. A new
sensor is initially assigned a total number of events which it will
detect from \numEventProtosDetected, and some set of attack
events. Each sensor is assigned at least one attack, plus
additional attacks up to its total number of events as consecutive
samples of a Bernoulli random variable with success probability
\sensorOverlap\ return 1.  After all sensor prototypes are
created and assigned attacks, we assign benign events to the sensors
to reach their total number of events.
%% \begin{itemize}
%% \item If we are given \numBenignEvents, then we create that total
%%   number of benign event prototypes, and distribute them among the
%%   sensors using a \sensorToBenignRatio\ which allows each sensor's
%%   total event count to be satisfied.
%% \item
%%   If we are given a \sensorToBenignRatio, then we create the
%%   necessary number of benign event prototypes to satisfy both
%%   \sensorToBenignRatio\ and each sensor's total event count.
%% \end{itemize}
We create enough benign event prototypes to satisfy
both \sensorToBenignRatio\ and each sensor's total event count, and
distribute them randomly among the sensors according to their total
event counts. 
% Finally we generate a network with the number of nodes
% determined by \numLocations.

%% (nondegenerate-simulate-step setting ...)
%%   > (simulate-step setting)
%%   > (events-for-location setting location)
%%     (reports-for-events setting location events)
% From the setting we can generate the ground truth aspect for an actual
% \mifd\ run: actual attack and benign events, plus the sensor reports
% which those events (or false positives) trigger. We proceed by
% considering each location of the network in turn.  We create the
% events which occur at each location by considering each attack and
% benign prototype. For each prototype, if a sample of a Bernoulli
% random variable whose success probability is the occurrence
% probability of the prototype is 1, then the ground truth aspect will
% contain a sensor report with that prototype and location.
% %
% For sensor reports at each location, we consider each sensor report
% prototype of the setting. If a prototype indicates that the sensor can
% detect one of the events which we have generated for that location,
% then we may have a report based on the events; otherwise we may have
% report because of a false positive.
% \begin{compactitem}
% \item If we have detectable events, then for each one we sample a
%   Bernoulli random variable whose success probability is the false
%   negative probability of the report prototype. If \emph{any} of these
%   samples are 0 --- that is, if for any event we do \emph{not} have a
%   false negative --- then the ground truth aspect will contain a
%   sensor report with that prototype and location.
% \item If there are no events at a location which could trigger a
%   report of that prototype, then we consult a Bernoulli random
%   variable whose success probability is the false positive probability
%   of the sensor report prototype; if it returns 1, then the ground
%   truth aspect will contain a sensor report with that prototype and
%   location.
% \end{compactitem}

\textbf{The experiments.}
% Our experiments probe the performance limits of our
% information fusion approach.  Our first two sets of tests examine how MIFD's
% performance degrades as events considered qualitatively different grow
% closer in likelihood.
% We then test how MIFD's performance degrades as the sensors it
% incorporates grow less and less discriminating.
% Finally, we examine how MIFD's sensor fusion helps base rate issues,
% and provide some analytic information.

\emph{Varying probabilities.}
Our first experiment examines how MIFD's performance degrades as
we progressively violate the assumption that different levels of
the stratified likelihood ranking qualitatively differ.
Specifically we consider a sequence of settings in which we assign
probabilities to each \tkappa,
always with
$P(\kappa=0)=0.5$ but varying the values for 1 and 2.
We test with \sensorToAttackRatio{}s of 2 and 3 to see how much corroboration is
necessary.  In the case of two sensors/attack we took \fpKappa=2, \benignKappa=2, \attackKappa=3
instead of the defaults above, in order to satisfy
Equation~\ref{eqn:coherence-ineq}.
We bring the probabilities corresponding to
the \tkappas closer and closer to see how MIFD's performance degrades.

\emph{Non point-value distributions.}
The above experiment still represents a considerable abstraction:  in general
the set of events to which we assign the same qualitative likelihood will not
all have the same probability.
Our second experiment examines how variation in the probabilities of the
events at the same qualitative likelihood affects the accuracy of qualitative
fusion.
To do so, we
define a second order probability distribution
for each \(\kappa\) ranking
using a \emph{beta distribution},
the Bayesian prior for Bernoulli
distributions.
The $\beta$ has two parameters, $\alpha$ and $\beta$.
The sum $\alpha+\beta$
corresponds to increasing the
``virtual sample size,'' and controls the variance.
% Beta distributions are appropriate for distributions over a finite
% domain such as the translations of these $\kappa$ values.
To assess the sensitivity of MIFD to variance in the probabilities,
we fix the mean values
of the distributions for $\kappa(0)$,
$\kappa(1)$ and $\kappa(2)$ at $0.5$, $0.01$ and $0.001$,
and increase the variance by decreasing the number of virtual samples.
% representing decreasing certainty in our parameter assignment.
% and vary the sample size $\nu=\alpha+\beta$, taking $\alpha=\mu\nu$,
% $\beta=(1-\mu)\nu$ for the translated distributions.  Higher sample
% sizes give smaller variance, and are thus closer to the single-value
% translations above; with smaller sample sizes the distributions will
% overlap, in the cases of our distributions for $\kappa(1)$ and
% $\kappa(2)$ with $\alpha<1$ and thus modes at $x=0$.

\emph{Sensor imprecision.}
We next explore how the performance of our techniques degrade as the
sensors encounter an increasing overlap of possible events in their
``field of view.'' 
For this experiment we fix the \sensorOverlap\ to be 0.9, and vary
\numEventProtosDetected. This value for \sensorOverlap is
significantly higher than in earlier runs, but we do not expect
performance in Setting~1 of this experiment to differ greatly from
Setting~1 of the earlier experiments because of the low initial values of
\numEventProtosDetected:  80\% of the time in this first run there will be only a single
event associated with a sensor, in which cases the \sensorOverlap\ is
not consulted at all.  With the higher
values from \numEventProtosDetected\ in subsequent settings, more
attacks will be associated with each sensor, increasing the opportunity
for \mifd\ to misdiagnose the true cause of sensor reports.

\emph{Base rate.}
Our fourth experiment examines how our performance
degrades
as the rate of true attacks goes down.
Such base rate problems, where even a high accuracy sensor can perform unacceptably
for very unlikely events, plague intrusion detection~\cite{Axelsson:1999:BFI:319709.319710}.
We examine these issues by reducing
the probability of generating attack events
(but not the probability of generating benign events). \hide{We do not actually change the
\attackKappa\ value used for the assessment step,
but instead introduce a
discrepancy between the model and the actual attack event probability.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 ids-fusion.tex
Initial import from SVN.

\section{Intrusion detection fusion}
\label{sec:ids-fusion}
% \longidses ({\idses}) are systems that sense intrusions in computer networks and
% hosts.  
{\ids} fusion % \footnote{In the computer security community, this process is often called ``\ids
% correlation,'' but we prefer ``fusion'' as it more accurately describes the
% process, and doesn't collide with statistical terminology.
% Indeed, in some cases the term ``\ids correlation'' is not used to describe an
% automated process, but rather a human-executed process of investigating an
% \ids-generated alert~\cite[e.g.]{drew14:_what_role_secur_event_correl_intrus_detec}.} 
is the
problem of creating a
coherent overall picture of network status out of reports from multiple
{\idses} scattered around a computer network. % we wish to defend.  
An \ids fusion system must answer two questions:
\begin{compactenum}
\item How do the reports issued by the \idses correspond to events? Multiple
  reports, from the same \ids or from different \idses,  might
  actually refer to the same event. This is the problem of \emph{clustering}
  reports into events.
\item Which hypothesized events do we accept as actual?  This is the problem of
  \emph{assessing} the likelihood of event hypotheses.
\end{compactenum}
In this paper we will focus on the second of these two questions, assuming a
solution to the first.  For discussions of the first question, see
\anoncite{goldman:09Scyllarus}.

Existing \idses are not designed to work together as part of a
suite of sensors.  Instead, each program generates a separate,
often voluminous, stream of reports; fusing them into a coherent 
overview is left as an exercise for the user. 
Ideally, network administrators would have a suite of different \idses active,
since different \ids approaches --- network (NIDS) vs.\
host-based (HIDS), signature vs.\ anomaly-based --- bring different
strengths and weaknesses.
%%
%% Hm, can we just get rid of the details NIDS/HIDS and
%% signature/anomaly comparisons, and keep just the takeaway? I don't
%% believe we refer to them after this. 
%%
Briefly, more accurate \idses have significantly higher software
administration burdens, and there is an inevitable tradeoff between
false positives and false negatives.
%%
%%Some systems aim to overcome the limitations of existing \idses by assembling 
%%a suite of
%%sensors.  This can be a very efficient way to overcome the problem
%%of false positives, but relies crucially on finding sensors that fail
%%relatively independently.  
%%
%% 
%% NIDS systems are easy to deploy; from only one
%% or two locations, they can watch all relevant traffic.  Unfortunately, they also have
%% substantial problems with false positives and false negatives, because they
%% don't ``know'' the meaning of the traffic they see.  Often they must draw
%% conclusions based only on information in packet headers (although some
%% systems do attempt to reconstruct higher level protocols from packet data),
%% and more extensive reasoning about traffic either would slow down the
%% network, or must ignore some packets altogether.
%% A HIDS may take into account much more information about the state
%% of a defended system, but installing HIDS pervasively involves deploying and
%% configuring much more software.  Typically HIDS are installed on only the most
%% critical assets.
%% 
%% Signature-based systems attempt to match patterns of known bad behavior.  Such
%% \idses are plagued by false negatives, both because of novel attacks for which
%% there are no patterns to match, and because of known %knowledgeable attackers take
%% countermeasures against signature identification (e.g.\ encrypting malware, or
%% permuting its structure).  Anomaly detectors are prone to false
%% positives: anomalies are often benignly anomalous,
%% not maliciously so.  Detectors often believe there is an
%% anomaly when none exists because computer usage patterns are
%% notoriously difficult to learn: they are often non-stationary, they vary
%% substantially from location-to-location, etc.
%% \draft{Add citations of papers on the challenges of anomaly detection.}

The most substantial challenge in managing \idses is the information overload
that they can impose.
Much of this overload comes from the high false positive rate.  This rate is
partly due to
inaccurate sensors.
Another problem is the ``base rate fallacy''~\cite{Axelsson:1999:BFI:319709.319710}:
in intrusion detection, as in
other applications where very unlikely events must be detected, even a seemingly
accurate sensor may yield too many false positives.  Axelsson gives the example
of a 99\% accurate test detecting a disease that strikes only 1/10,000
patients.  The probability of having the disease given a positive test result is
only 1\%, although the probability of the sensor labeling correctly is 99\%.
{\ids} users regularly either ignore or partially disable them, unable to absorb
massive streams of false alarms:
``Users attending an 'ABCs of IDS' event at London's City University yesterday
  said more the 80 per cent of the alerts they received were false, with one
  citing 60 alerts he had received about non-existent problems that morning at
  0300''~\cite{Register}.
Recently, Target is reported to have ignored warnings about the
data breach that resulted in theft of millions of credit
cards~\cite{TargetIgnoredIDS}:
``They are bombarded with alerts. They get so many that they just don't respond
to everything''~\cite{finkle14:_target}.
Figure
\ref{fig:workloadReduction} gives a sense of the gravity of this
problem, showing how our earlier system was able to winnow
the 
flow of reports in a small corporate network~\anoncite{goldman:09Scyllarus}.
Data reduction, in addition to building a comprehensive overview
and overcoming false positives and negatives, must be a goal of \ids fusion.
% \begin{figure*}[t]
%   \begin{tabular}{c}
%     \ifthenelse{\boolean{blackAndWhite}}{%
%       \includegraphics[width=.7\textwidth]{figures/funnel-bw.pdf} }{%
%     \includegraphics[width=.7\textwidth]{figures/funnel.pdf}} \\
%     \ifthenelse{\boolean{blackAndWhite}}{%
%         \includegraphics[width=.7\textwidth]{figures/funnel-graph-bw} }{%
%             \includegraphics[width=.7\textwidth]{figures/funnel-graph} } 
%   \end{tabular}
%   \centering
%   \caption{Scyllarus workload reduction~\anoncite{goldman:09Scyllarus}.}
%   \label{fig:workloadReduction}
% \end{figure*}
\begin{figure}[t]
    \ifthenelse{\boolean{blackAndWhite}}{%
      \includegraphics[width=.45\textwidth]{figures/funnel-dup-cropped.pdf} }{%
    \includegraphics[width=.45\textwidth]{figures/funnel-dup-cropped.pdf}} 
  \centering
  \caption{Scyllarus workload reduction~\anoncite{goldman:09Scyllarus}.}
  \label{fig:workloadReduction}
\end{figure}

% For example:
% \begin{quotation}
%   Users attending an `ABCs of IDS' event \ldots 
% %at London's City University yesterday 
% said more the 80 per cent of the alerts they received were
% false, with one citing 60 alerts he had received \ldots 
% %about non-existent problems 
% that morning at 0300....
% %The concern is that 
% [A]dopters \ldots
% %of the technology 
% will \ldots %fail to maintain it or 
% simply leave it to gather dust as overworked admins get bombarded with false alarms.~\cite{Register}
% \end{quotation}

% Often the most damning weakness of an IDS is a high false positive rate. 
% In general, with any
% sensor, one must pay in false positives for whatever is gained in
% sensitivity.
%\footnote{Receiver operating characteristic (ROC) 
%curves provide a way of assessing how
%well a particular sensor can provide true detections for a particular
%cost in false positives.  The ROC curve is the curve one gets 
%when plotting $p(\mbox{hit})$ vs $p(\mbox{false positive})$ as 
%the detector's threshold is swept from its most conservative
%to most liberal settings~\cite{MacMillanCreelman:91}.}

% Data reduction, a comprehensive overview, and overcoming false positives and
% false negatives are the goals of \ids fusion.
% Such systems aim to overcome the limitations of existing \idses by assembling 
% a suite of
% sensors.  This can be a very efficient way to overcome the problem
% of false positives, as long as we can find sensors that fail
% relatively independently.

% A shared framework like ours can also provide protection against
% {\em systematic\/} false positives.  For example, in an early installation of
% \correlator{}, we worked with (an earlier version of) the IDS {\tt
% snort}, which contained rules for detecting IP sweeps.  We found that,
% in sites where \longnav\ were run, this rule was tripped by the normal operation of \nav .
% An automatic update server for \nav\ periodically scanned the network,
% looking for hosts running the \nav\ update client.  
% Such clients listened on
% \navport.\footnote{We are not singling out {\tt snort} for
% criticism.  Our experience has been with {\tt snort}, but
% virtually all network-based \ids es would exhibit this kind of behavior.}

% The ordinary solution to this kind of problem is to edit the \ids\ rule to
% keep it from firing in this circumstance, for example by ignoring IP
% sweeps that hit only \navport .
% This approach is unsatisfactory for two reasons.  First, in sites
% where multiple IDSes are run, this means that we must separately
% update the configuration of each individual IDS.  
% Second, it allows a clever attacker free rein with traffic on this
% port.

% On the other hand, if we have a central repository of this kind of
% information, we get two corresponding advantages:
% First, we have only a single point of update.  Instead of fixing each
% individual sensor, we reconfigure so that all of the reports
% corresponding to this class of false positives will be filtered out.
% Second, we can collect additional information, for example, logs of
% the \nav\ server, that allow us to distinguish between true \nav\
% events, and transmissions from a clever attacker that are disguised as
% \nav\ transmissions.

\draft{Challenges of IDS evaluation}

\paragraph{Related work.}
SecurityFocus has developed the
Attack Registry and Intelligence Service~\shortcite[ARIS]{ARIS}.
Its extractor consolidates {\ids} reports from four different
{\idses} into XML, and presents them in an incident
console.  However it makes no attempts to fuse the reports, nor weigh
evidence for and against them.
MetaSTAT is a fusion system built on a set of STAT-based
{\idses}~\cite{vigna01:designing}.  
STAT is signature-based, and detects events by matching
against extended finite-state event models.  MetaSTAT uses
finite-state models of across-sensor events to consume events
from lower-level sensors at a higher level,
but does not attempt to judge the plausibility of different
events.
EMERALD/eBayes~\cite{raid2001-pac} fusion is the most similar to our systems.
Their sensors are Bayes net-based, and the correlation approach
allows ``upstream'' sensors to adjust the priors on ``downstream''
sensors.
Its fusion is limited to clustering alerts that meet a
similarity criterion; they do not have models of high-level events like
ours (see below).  To the best of our knowledge, they do not
address the difficult issues of acquiring probability parameters for \ids fusion.
The Prelude IDS's Correlator allows users to analyze reports sent to Prelude 
from compatible IDSes~\cite{PreludeCor} following user-provided rules.  Its function is closest to
our clustering subsystem, but knowledge resides in stateful 
rules instead of an ontology of attacks. 
The Arcsight Enterprise Security Manager~\cite{ArcsightEsm},
also ties correlated IDS reports to an installation's security goals
and vulnerability information.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 intro.tex
Initial import from SVN.

\section{Introduction}
\label{sec:introduction}

We experimentally analyze the accuracy of the \zplus qualitative
probability scheme when used for diagnosis and information fusion.
In cyber defense, an ``intrusion detection system (IDS) is a device or
software application that monitors network or system activities for malicious
activities or policy violations''~\cite{wiki:ids}.
In previous work we developed a technique for IDS
fusion, deployed in the Scyllarus system~\anoncite{goldman:09Scyllarus}, and its
successor MIFD~\anoncite{stratus:saso12}.
These systems fuse together reports from multiple, heterogeneous IDSes,
hypothesizing underlying events to explain those reports, and assessing the events'
likelihood, to detect cyber attacks.
Their likelihood assessment is based on \zplus, a qualitative
abstraction of probability theory~\cite{Goldszmidt:96}.
Scyllarus has been extensively tested in real networks, using both
real and synthetic data, and has shown its ability to accurately fuse reports
from extremely noisy sensors. 
We also
drastically reduce false alarm rates, the bane of intrusion
detection systems, reducing the flood of incoming reports by multiple
orders of magnitude.
Unfortunately, it is difficult to draw crisp, \emph{general} conclusions based on
field evaluations alone.
We need to demonstrate that the results are due to features of the algorithm,
not simply artifacts of details of the test network, traffic, and
attacks.  This problem is particularly acute in the area of intrusion detection,
as we explain below.
% This is a general problem of evaluating IDSes, which we discuss
% further below.
%

In this paper, we analyze the underlying reasoning machinery to
complement earlier field studies.
%
% In particular we explore the suitability of qualitative
% probabilistic techniques --- and \zplus in particular ---
% for information fusion, especially in areas where
% standard probabilistic methods cannot be applied because statistics are not
% available, and machine learning techniques are not appropriate.
\zplus models uncertain phenomena as falling into a small set of
\emph{qualitatively distinct} levels of likelihood, similar to the way that
big-$O$ methods abstract computational effort.
% In our experiments, we show that our abstraction supports accurate event
% detection, and explore the robustness of our techniques as
% the qualitative abstraction fits the world less well.
To use the big-$O$ analogy again, we check to make sure that our results degrade
gracefully as the orders of magnitude become less important compared to the
constant factors.
% For example, we examine how our qualitative approach degrades when
% modeling probability distributions in which the probabilities corresponding to the
% qualitative strata get closer together.

%
% %%% Can we remove this? It's a bit early for future work, so maybe
% %%% push to back if necessary and space allows?
%
%Additionally, our original work was in the context of an open system that fused
%a set of ``take them or leave them'' \idses.  We are moving to incorporate our
%fusion system in an integrated system that will perform computer network
%defense.  In this new framework, we have the opportunity to make decisions about
%what sorts of sensors to deploy, possibly developing new sensors in the process,
%and where to deploy these sensors.  The results here will help inform such
%decisions.

The experimental results we report show that \zplus{}'s accuracy degrades
gracefully as the qualitative abstraction fits less and less well. We also show
that the accuracy of our system degrades gracefully with decreasing sensor
precision. Finally, MIFD is accurate even when detecting very rare events,
not only when sensors fail independently, but also in
the face of correlated false positives.
These results confirm the results of our
earlier field tests, and help explain why the
qualitative scheme works so well.

% Our experimental analyses use simulated sensors and events,
% allowing us to precisely control stimuli to the system,
% and so allow us to draw general conclusions about the applicability of our
% approach.
% %, to identify when it will behave well or poorly, show that it degrades
% % gracefully, rather than suddenly, etc.
% The experiments are very encouraging, confirm the results of our
% earlier field tests, and help to explain why it is that the
% qualitative scheme has been found to work so well.
% % They show that under very weak assumptions, 
% We show that
% MIFD's sensor fusion approach provides good detection rates with a low
% rate of false positives even when the underlying assumption that the
% likelihoods of events do qualitatively differ is violated.  The
% experiments further show that MIFD functions correctly even when the
% connection of sensors to events is ambiguous.  MIFD achieves
% acceptable false positive rates even when detecting very rare events,
% and can do so not only when sensors fail independently, but also in
% the face of correlated false positives (represented as \emph{benign}
% events).

Our results are generally
relevant to qualitative probabilistic reasoning for
information and sensor fusion.
Our fusion problems are modeled as problems of causal explanation, or abduction:
what are the events most likely to have caused the observations (the IDS
reports) given certain causal relations?
Therefore our results are also of interest to researchers in diagnosis.
Qualitative probability systems like $Z+$ offer an attractive middle point between
purely disjunctive reasoning in diagnosis, and full probabilistic reasoning.

To the best of our knowledge, ours is the only empirical work to explore the
\emph{accuracy} of reasoning with \zplus.
Of about 250 works citing Goldszmidt and Pearl's 1996
work~\cite{google-scholar-query}, none of the few which detail
applications of \zplus\ reasoning conduct such investigations.
%
Minock and Kraus~\shortcite{minock-kraus-LIAI02} investigate the
\emph{efficiency} of an implementation of \zplus,
\hide{restricted to specific classes of
Horn theories,}
but not its accuracy.
% Minock and Kraus compile queries over Horn and q-Horn clauses to an
% implementation of \zplus, but they report only ``initial performance
% results'' for partition construction to illustrate the tractability of
% their approach~\shortcite{minock-kraus-LIAI02}.
%
% Fujimoto and Matsuzawa use qualitative rankings similar to \zplus\ to
% categorize subjective evaluations in
% text~\shortcite{fujimoto-matsuzawa-newgencomp99,fujimoto-WIIAT2010};
% their system predicts reactions to advertising messages.  Fujimoto's
% experiment~\shortcite{fujimoto-WIIAT2010} is directed towards a
% regression analysis, examining prediction accuracy on a set of survey
% responses.
%
% We do not question the correctness of those experiments, but
% the substitution of qualitative probabilities for actual values should
% not be made without confirming the appropriateness of the
% approximation for the particular domain.  We believe that the
% experiments we report here are the first such rigorous investigation
% of the suitability of \zplus\ for a specific application.

In the next section we introduce the problem of \ids fusion
and its challenges.  Then we describe our approach to the
problem, as implemented in the Scyllarus and MIFD systems.  We describe our
experimental designs, present the results and conclude with some proposals for
future work.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 results.tex
Initial import from SVN.

\section{Results}
\label{sec:results}

\begin{figure}
  %% \labresults{Single-point distributions}%
  %%            {figures/point-distributions-cropped.pdf}
  %% \begin{center}
  %%   $\begin{array}{r|c@{~}c|}
  %%   \multicolumn{1}{c}{}
  %%   & P(\kappa=1)
  %%   & \multicolumn{1}{c}{P(\kappa=2)}
  %%   \\ \cline{2-3} 1 & 0.005 & 0.0005
  %%   \\ 2 & 0.01  & 0.001
  %%   \\ 3 & 0.05  & 0.005
  %%   \\ 4 & 0.1   & 0.01
  %%   \\ \cline{2-3}
  %% \end{array}$
  %%   \hspace{\fill}
  %%   $\begin{array}{r|c@{~}c|}
  %%   \multicolumn{1}{c}{}
  %%   & P(\kappa=1)
  %%   & \multicolumn{1}{c}{P(\kappa=2)}
  %%   \\ \cline{2-3} 5 & 0.25  & 0.125
  %%   \\ 6 & 0.33333 & 0.22222
  %%   \\ 7 & 0.375 & 0.28125
  %%   \\ 8 & 0.46875 & 0.439453125
  %%   \\ \cline{2-3}
  %% \end{array}$
  %% \end{center}
  \input{figures/point-distributions-table}
  \caption{Precision (P) and recall (R) percentages from the \textbf{varying
      probabilities} experiment.  The second result set varies
    the initial sensor-to-attack ratio (s2a) from 3 to 2.
    %%
    %% The lower key relates the labels on the x axis to assignments
    %% of probabilities to kappas. $P(\kappa = 0) = 0.5$ for all.
  }
  \label{fig:vary-kappa-results}
\end{figure}

%% IN JOURNAL VERSION --- smoothed vs. unsmoothed results; see
%% also commented-out lines in figures/vary-kappas-beta-unified.gip
\begin{figure}
    \labresults{(i) Beta distributions}{figures/vary-kappas-beta-unified-cropped.pdf}
%%  \caption{Precision and recall percentages for the \textbf{non
%%      point-value distribution} experiment. 
%%    %The mean values of the
%%    %distributions translating $\kappa(0)$, $\kappa(1)$ and $\kappa(2)$
%%    %are respectively $0.5$, $0.01$ and $0.001$; the sample sizes
%%    %$\alpha+\beta$ vary as shown.
%%  }
%%  \label{fig:vary-beta-results}
%%\end{figure}
%%\begin{figure}
\\  \labresultsT{(ii) Sensor imprecision}{figures/crowding-sensors-cropped.pdf}
%%  \caption[Precision and recall percentages of the \textbf{crowding
%%      sensors} experiment.]%
%%          {Precision and recall percentages of the
%%  \textbf{crowding sensors} experiment.}
%%  \label{fig:vary-crowding-results}
%%\end{figure}
%%\begin{figure}
\\  \labresultsT{(iii) Base rate}{figures/base-rate-cropped.pdf}
%%  \caption{Precision and recall percentages of the \textbf{base rate}
%%    experiment.}
%%  \label{fig:vary-baserate-results}
\caption{Precision and recall percentages for the \textbf{beta
    distribution}, \textbf{sensor imprecision} and \textbf{base rate}
  experiments. }
  \label{fig:vary-beta-results}
  \label{fig:vary-crowding-results}
  \label{fig:vary-baserate-results}
\end{figure}

\begin{figure*}
  \centering
  \textsf{Three sensors}\hspace{40mm}\textsf{Two sensors}
  \hspace{38mm}\textsf{Single sensor} 
  \\
  \includegraphics[width=\textwidth]{figures/consolidated-base-rate-cropped.pdf}
  \caption[Base rate problems and sensor fusion.]
%%
    {Graphs of quantitative false positive probabilities (on the
      z-axis) as function of the probability of an individual sensor
      giving a false positive reading (on the x-axis, running from the
      origin to the lower right on each graph) and attack probability
      (on the y-axis).  Specifically, each graph plots
      $z=P(\mathrm{no~attack}|\mathrm{alarm~raised})=\frac{x^s(1-y)}{y+x^s(1-y)}$
      for the given number of sensors $s$. }
    %%{Base rate problems and sensor fusion.  The graphs show the false
    %%  positive probabilities (on the z-axis) as a function of the
    %%  false alarm probability (on the x-axis, running from the origin
    %%  to the lower right on each graph) and attack probability (on the
    %%  y-axis).  The diagonal contour lines project onto the xy-plane
    %%  where the surface intersects z=0.25, z=0.5 and z=0.75.}
  \label{fig:base-rate-plots}
\end{figure*}

\emph{Varying probabilities.}
Figure~\ref{fig:vary-kappa-results} shows \mifd's performance as the
stratified likelihood rankings become less distinct.
We see that MIFD's performance degrades gracefully as the probabilities
corresponding to $\kappa(0), \kappa(1)$ and $\kappa(2)$ approach each other.
Although the
theoretical basis of the qualitative probability levels is
infinitesimal, \mifd\ performs reasonably for $P(\kappa=1)$ and
$P(\kappa=2)$ taken as high as 0.05 and 0.005 respectively.
\mifd\ also degrades gracefully as its input sensors become
less discriminating, simulated by raising the sensor-to-attack ratio
from 2 to 3.  We had expected \mifd\ to be more brittle under
sensor-to-attack ratio 2, but this was only partly true;
precision declines more quickly with ratio 2 than with ratio 3.
Recall declines less quickly, because the corroboration of several
sensors
is more relevant to suppressing spurious additional event diagnoses.

\emph{Non point-value distributions.}
Figure~\ref{fig:vary-beta-results}\,(i) shows \mifd's performance when we
interpret the qualitative likelihood levels as various beta
distributions.
The results confirm those of the previous experiment, and show that MIFD is not sensitive
to variance in the probabilities of events at the same qualitative surprise level.

\emph{Sensor imprecision.}
Figure~\ref{fig:vary-crowding-results}\,(ii) shows how \mifd's accuracy
declines when sensors correspond to more than one event.  
Recall and precision stay well above 90\% for sensors responding to
two or fewer distinct events. Unsurprisingly, precision declines sharply as
sensors respond to three or more events.  This decline is not
particularly serious in practice: actual \ids\ sensors rarely detect
large numbers of events: if they do we introduce new
abstractions.

\emph{Base rate.} \label{sec:base-rate}
Figure \ref{fig:base-rate-plots} shows the challenge of base rates, and how MIFD
addresses it through sensor fusion.  
In the rightmost plot, we see
that the
false positive probability is extremely high with only one sensor, even a very
accurate
one.  However, requiring corroboration from two or three sensors,
as MIFD does, reduces the false alarm rate substantially.
Note that this requires the sensors to fail independently.  If sensors'
false positives are correlated, then corroboration can fail to
lower the false positive rate.  It was
to handle such correlated failures that we introduced benign events
into Scyllarus and MIFD.

Figure~\ref{fig:vary-baserate-results}\,(iii) shows how \mifd's false
positive rate rises with event rarity. The high recall shows that
\mifd\ does find the attacks (the bounce up at the right is because of the
extremely low sample size), but declining precision is from mistakenly
hypothesizing additional attacks.  Precision declined unacceptably
below 90\% for smaller attack probabilities
than shown on this chart.

% LATER: What accounts for the oddball results here?

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 scyllarus-and-mifd.tex
Initial import from SVN.

\section{The Scyllarus and MIFD systems}
\label{sec:mifd}

Our original system Scyllarus was an open system which performed IDS
fusion on a diverse set of third party \idses~\anoncite{goldman:discex2001,goldman:discex2001}.
Its successor
MIFD (Model-based Intrusion Fusion and Detection),
enhances Scyllarus as part of the STRATUS system's
integration of IDS fusion into a comprehensive suite for cyber defense
\anoncite{thayer13:_compar_strat_tactic_respon_cyber_threat}.
For simplicity we refer to our \ids fusion system as ``MIFD,'' but
except where we specify otherwise, our account applies to Scyllarus as well.

The first tasks performed by MIFD are
\emph{collection} and \emph{translation}: IDS reports
must be collected and presented to MIFD, and
must be translated into a common
sensor report data structure for processing.
We will not discuss these less interesting
preliminaries, except to mention that they
involve a substantial amount of data rectification.  There is no widely-accepted
common data format (the proposed IDMEF standard~\cite{IDMEF} has not caught on), nor
is there an accepted taxonomy for events or event roles (e.g., the widely used ``source''
and ``destination'' are not consistently applied).
These problems are exacerbated by the fact that intrusion detection ``systems''
are often simply a casually-assembled anthology of rules by different authors running in the same
framework.

Core fusion inference begins with
\textbf{Clustering:} From the
sensor reports, MIFD generates \emph{event hypotheses} to represent the
underlying events of interest that may have caused the sensor reports.  A
sensor report that provides evidence for an event hypothesis is a
\emph{supporter}.
The clustering process constructs a
directed graph of event hypothesis and sensor report nodes forming a
Bayesian belief network.

Figure \ref{fig:bnets}(a) shows the Bayes net generated by MIFD for
the very simple case of a single event hypothesized to be the cause of
a number of sensor reports.  In general sensor reports will be
ambiguous, each supporting several different event hypotheses.
Figure~\ref{fig:bnets}(b) shows a more complex example where sensor reports
might be caused by two different underlying events.
%% DONE: cut the following sentence because its meaning wasn't
%% clear.... I think what's meant is something like "label" rather than
%% "measurement," but I couldn't figure out how to clarify it. --- How
%% about this?
% The sensors themselves are simply either on or off, without any additional
% outputs which would distinguish hypotheses.

\ids rules detect indirect features of intrusions, but they typically
report only the intrusion that their designer believes has caused the feature.
This creates a number of problems, first and foremost that the
sensors have high false positive rates.
For example, a local software update server
may look like an attacker performing reconnaissance.
What's worse, some of the false positives are correlated: the normal
activity of a print server may be mis-detected as a scan by several
different sensors.
To handle simple sensor failures, MIFD associates with each sensor a false
positive likelihood.
But to address the correlated false positives such as the scanning example, we
introduce benign events, that can compete with attack and (simple) false
positive explanations for results.\footnote{Where possible, we try to introduce sensors that
can specifically detect the benign events. To return to our earlier example, we
might introduce a ``print server sensor'' to help disambiguate detections. We do
not address this issue in our experiments.}
% FIXME: Add a figure 2(c) -- or separately numbered figure -- where we have a
% benign event, explaining two of the sensor reports, as well as the two attack
% events. 
Another problem with indirect sensing is that the sensors are often imprecise:
the features that they detect often occur not only in the absence of the
advertised intrusion, but in the presence of \emph{other} intrusions with
similar characteristics.

MIFD also has \emph{complex} events that have component sub-events, such as a
phishing attack with component events phishing email and drive-by-download.
Our current experiments do not address the role of complex events, so we will
not discuss them further.

\begin{figure}
  \centering
  \begin{tabular}{c@{\hspace{1cm}}c}
  \includegraphics[width=.2\columnwidth]{figures/bnets1-cropped.pdf} &
  \includegraphics[width=.267\columnwidth]{figures/bnets2-cropped.pdf} 
  \\ (a) & (b)
  \end{tabular}
  \caption{Bayes networks generated by MIFD.}
  \label{fig:bnets}
\end{figure}

The final inference step is \textbf{Assessment:} MIFD assigns a degree of belief
to each of the event hypotheses.
MIFD does this by performing Bayesian updating on the belief network constructed
by the clustering process. 
%% As we discuss below, MIFD does this using the \zplus qualitative probability
%% scheme rather than standard probability theory.
%%
Scyllarus and MIFD are intended as long-running online services, not batch processes.
Scyllarus has run for months at a time.
One major requirement for such long-term operation is that the
inference steps must be performed cyclically.  MIFD repeatedly reads
reports and clusters them, periodically interrupting that
cycle to perform assessment.
%% FIXED Although this next sentence is a true fact, I'm not sure it
%% belongs here --- it makes assessment efficient, but that's not per
%% se an inherent requirement for long-term operation.
% Moreover, in general the Bayesian belief network
% is not fully connected, so for efficiency MIFD assesses each connected
% component individually.
% \footnote{%
%   An additional essential requirement for long-term operation, not
%   relevant to the remainder of this presentation, is some form of
%   garbage collection: old event hypotheses and sensor reports
%   guaranteed no longer to be relevant to subsequent report must be
%   removed from working memory.}

\input{zplus}

% Not sure that the reader cares about the data model, yet.  Maybe; maybe not.
\hide{%
\mifd\ operates under the following data model:  It receives
\emph{sensor reports} and generates \emph{event
  hypotheses}. Hypotheses include a \emph{likelihood} slot, which
\mifd\ may update as it receives additional reports, plus a list of
\emph{supporters} which may be either sensor reports or other
hypotheses. Both sensor reports and hypotheses have a list of
\emph{supports}, inverse to the \emph{supporters} relationship.
Both reports and hypotheses are associated with \emph{prototype}
objects which contain information common to all instances sharing a
prototype. In particular:
\begin{compactitem}
\item Sensor report prototypes include a slot for the (natural
  number-valued) qualitative probability of false positives for the
  report.
\item Event prototypes include a slot for the qualitative
  probability of the event occurring.
\end{compactitem}
The relationship between reports and events is modeled to \mifd\ as
\emph{hypothesis matchers}, which pair a report prototype and an event
prototype. Each instance where a particular sort of event might
trigger a particular sort of report is modeled as a separate
hypothesis matcher, so there may be several hypothesis matchers
pertaining to each sensor report prototype.  In an initializing step,
\mifd\ creates a table of hypothesis matchers indexed by sensor report
prototypes.

\mifd\ provides two operations, \emph{clustering} and
\emph{assessing}.  Clustering occurs whenever \mifd\ receives a new
sensor report.  For each of the event prototypes paired with the new
report's prototype by some hypothesis matcher,
\begin{compactitem}
\item If there are existing hypotheses with that event prototype, then
  \mifd\ adds the new report as a supporter of these hypotheses.
\item If there are no existing hypotheses with that event prototype,
  then \mifd\ creates a new hypothesis with the new report as its sole
  supporter.
\end{compactitem}

Assessment occurs whenever \mifd\ is required to produce updates to
hypothesis likelihood. Each sensor report could individually justify
possibly many events, but in assessment \mifd\ distinguishes more
likely hypothesized events from less likely ones, taking into account
all reports and the prior probabilities of hypothesized events and
report false positives.  We can first decompose this task by
considering the graph formed by the \emph{supporters} relationship on
reports and hypotheses, and consider the connected subgraphs in
turn. Moreover, we need consider only those subgraphs containing a
sensor report received since the last assessment.

To distinguish the more and less likely hypothesized events in each
connected subgraph, \mifd\ uses an assumption-based truth maintenance
system (ATMS). An ATMS is a directed, acyclic hypergraph in which
nodes represent possibly-true statements, and hyperedges from $N$
nodes to a single node represent an inference from the $N$ statements
concluding the single statement~\cite{Forbus+DeKleer:1993}. The
initial nodes of the ATMS hypergraph are called \emph{assumptions}; in
\mifd's use of ATMSs there are three forms of assumption:
\begin{compactitem}
\item ``Event $X$ did happen.''
\item ``Event $Y$ did not happen.''
\item ``Sensor report $Z$ was a false positive.''
\end{compactitem}
Corresponding pairs of the first two forms are marked in the ATMS as
\emph{nogood}.  Node representing that an event occurred can be
justified either the corresponding event assumption, or by other
events which can be generalized or collected together; nodes
representing the receipt of a sensor report can be justified either by
an event's node, or by a false positive and the negation of an event
happening.

As an ATMS's hypergraph is constructed, the ATMS algorithm maintains
the list of minimal consistent assumption sets that justify each node.
\mifd\ is interested in the most likely consistent assumption sets
satisfying the sensor report nodes, that is, those which have the
lowest total qualitative probability rank.  In principal \mifd\ could
simply add another ``goal'' node justified by the sensor report nodes,
use the ATMS algorithm to maintain its assumption sets, and find the
most likely ones. But in practice this is not a feasible approach;
finding all such consistent sets simply overwhelms the ATMS algorithm.
Instead \mifd\ performs an A* search to find the most likely unions of
assumption sets from each of the sensor node labels.

Finally, to determine the likelihood of each hypothesized event $E$,
\mifd\ looks for the presence of the assumptions that $E$ did or did
not happen in the most likely assumption sets. If only the positive
assumption occurs in these sets, then the hypothesis is deemed
\emph{likely}, the highest of the three output qualitative probability
values; if only the negative assumption occurs, then \emph{unlikely},
the lowest. If both positive and negative assumptions occur, then the
hypothesis is given a third, middle value \emph{plausible}.}

% Because the domain does not afford us access to good statistics, we do not use
% conventional Bayesian reasoning.  Instead, we use a qualitative abstraction of
% probabilistic reasoning, very similar to the big-O scheme familiar to computer
% scientists, \emph{\zplus}~\cite{Goldszmidt:92,Goldszmidt:92a,Goldszmidt:96}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 zplus.tex
Initial import from SVN.

% \subsection{System Z+}
% \label{sec:zplus}

MIFD
follows Pearl's exhortation~\shortcite{Pearlbook} that the core of probability theory is
the patterns of inference it enables (explaining away, combining forward causal
inference and evidential reasoning, etc.), rather than precise numerical
calculations.  He recommends these patterns be used even when precise numbers are
not available, and that is what MIFD does.  As outlined above, MIFD treats
\ids fusion as an abductive problem, formalized using Bayes nets.
But precise values for the parameters of these Bayes nets ---
probabilities of attacks, false positives, false negatives, etc. ---
are \emph{never} available to us: populations of
attacks and attackers change, different networks have little in common with each other, distributions are
non-stationary, etc.  Moreover these difficulties are not solvable via machine learning:
Sommer and Paxson~\shortcite{Sommer10outsidethe} discuss the challenges \ids pose to ML.

Goldszmidt and
Pearl's \zplus{}~\shortcite{Goldszmidt:96},
shares the
basic structure of normal probability theory, but abstracts the actual
probabilities.
Events have a natural number rank \tkappa{}
corresponding to their degree of surprise (so a rank of one is more surprising
than zero). The semantics of this scheme comes from a set of probability
distributions in which the probabilities are polynomials over some infinitesimal
$\epsilon$. In this scheme, the \tkappa{} corresponds to the exponent
of the polynomial's leading term.
As we mentioned earlier, \zplus is similar to the ``big-$O$'' method used
in analysis of algorithms.
This semantics gives \zplus{} a
qualitative flavor by providing a ``ladder'' of events of qualitatively different
orders of likelihood.
Note that \zplus is very different from Wellman's qualitative probability
networks~\cite{Wellman:90}, whose abstraction is based on direction of change, rather than order
of magnitude relations ``at rest,'' as it were.
Wellman's approach is more relevant when choosing interventions (as in his work
on medical diagnosis and treatment), and less for fusion.

\hide{
The \assessor{} must combine the judgments of a wide variety of intrusion
detection systems (and potentially other relevant information sources), that use
widely varying sources of information and algorithms. Further, in general we
will not have access to the internals of these sensors.  In such an environment,
it is not realistic to expect good models of the response of these sensors; in
particular, exact measures of P(sensor response|event) are not available.  There
have been some attempts to investigate sensor response (e.g., the studies
conducted by Lincoln Labs~\cite{lippmann00:eval}), but the results seem heavily dependent on the
context in which the sensors are deployed.

The issue of prior probabilities also militates against the use of exact
probabilities.  In order to use an exact Bayesian method, we would need not only
the detection probability, $P(\mathit{sensor\ response}|\mathit{event})$, and
the false alarm probability, $P(\mathit{sensor\ response}|\neg \mathit{event})$,
but also $P(\mathit{event})$, a measure of the prior
probabilities of the events that interest us, in this case the attacks and the
benign events that can cause false positives.  Even in the most constrained
environments, the probabilities of the various attacks, are unlikely to be
available to us,  and the Scyllarus system is designed for application across a
wide variety of enterprises.  Further, the probability distributions for benign
events are likely to be of odd forms (e.g., one's own network-mapping software
runs at particular times of the day).  So our solution must tolerate vague
measures of likelihood.

Finally, in this domain, as with most practical applications of probabilistic
updating, the effect of the evidence will \emph{usually} overwhelm the effect of the prior
likelihoods(e.g., \cite{pradhan:96}). So inexactitude in the quantities specified will not matter to our
final conclusions.}

Under \zplus\ we may apply the normal operation
of probability theory --- conditionalization, Bayes' law, etc. --- but
the arithmetic operations we use for them change.  Rather than multiplying
probabilities, we add degrees of surprise; rather than adding
probabilities, we take the minimum surprise.  Goldszmidt and
Pearl~\shortcite{Goldszmidt:96} provide the following substitutions
between the quantitative probability function $P$ and its qualitative
counterpart $\kappa$ for formulas $\phi$ and $\psi$, and primitive events, $e$.
\begin{displaymath}
  \small
  \begin{array}{|l|l|}
    \hline
  P(\psi) = \sum_{e \in \psi} P(e) & \kappa(\psi) =
  \operatorname{min}_{e \in \psi} \kappa(\phi) \\ \hline
  P(\psi) + P(\neg \psi) = 1 & \kappa(\psi) = 0 \vee \kappa(\neg \psi) = 0
  \\ \hline
  P(\psi | \phi) = P(\psi \wedge \phi) / P(\phi) &
  \kappa(\psi | \phi) = \kappa(\psi \wedge \phi) - \kappa(\phi)
  \\ \hline
\end{array}
\end{displaymath}
\noindent
In a Bayes network, with conditionally independent $\omega$ and
$\phi$ we further have $\kappa(\omega \wedge \phi) = \kappa(\omega) + \kappa(\phi)$.
\hide{Instead of the probability
of an event being the sum of the probabilities of the primitive outcomes that
make up that event, the degree of surprise of an event is the minimum of the
degrees of surprise of the primitive outcomes that make it up. Instead of having
the probabilities of mutually exclusive and exhaustive events sum to one, at
least one of a set of mutually exclusive and exhaustive events must be
unsurprising. Finally, we have an analog of Bayes' law in which the normalizing
operation consists of subtraction rather than division.
}

\hide{We used Bayesian networks to help us in modeling and solving the correlation problem. Bayesian networks are 
ways of graphically capturing probabilistic reasoning. They are useful in expert systems because they 
simplify knowledge acquisition and, by capturing (conditional) independences, simplify computation 
(Pearl, 1988). In particular, in the domain of intrusion detection, Bayes nets help us capture several 
important patterns or probabilistic reasoning:
\begin{itemize}
\item	Reasoning based on evidence merging; 
\item	``Explaining away'' reports by alternative explanations.  E.g., if a benign event accounts for a number 
of reports, those reports will be explained away, and no longer provide support for more alarming 
hypotheses.

\item	Abstraction reasoning that employs the sub\-class/su\-per\-class relationships in the event dictionary.

\item	Part/whole reasoning, to recognize complex composite events.

\item	Distinguishing between judgments that are based on different sensor bases and those that use the same 
sensor.  This helps us distinguish between cases when two sensors provide support for each other and 
when we simply have redundant reports (e.g., two network intrusion detection systems using exactly 
the same algorithm that see the same traffic, at two different points).
\end{itemize}
A Bayesian network is a directed, acyclic graph (DAG) depicting a set of random variables. Edges between 
nodes in the DAG represent causal influences. Using a Bayesian network, we can capture a joint 
distribution factorized into unconditional probabilities for root nodes and conditional probability tables for 
non-root nodes. The conditional probability tables contain probability distributions for the child nodes, 
conditioned on all the values of their parents.
}

There are a number of efficient algorithms for finding the posterior distributions of Bayesian networks, 
conditional on observations of some of the random variables. These algorithms may readily be adapted to 
provide posterior $\kappa$ rankings instead of probabilities.
MIFD translates Bayes networks into an
%% DONE Charniak:88 is Charniak and Goldman --- withhold this one too? No --
%% it's not obviously tied to this paper, so we don't need to worry about it
%% breaching anonymity. [2014/08/24:rpg]
ATMS~\cite{Forbus+DeKleer:1993}; see \cite{Charniak:88,Poole:93,provan-ijcai89}
and \anoncite{goldman:09Scyllarus} for this encoding.
This implementation is not especially efficient, but we have found
it is I/O which dominates runtime, and we apply several optimizations to
handle common special cases.
To date (somewhat to our disappointment!) optimizing the general \zplus
inference method has not been key to system performance.
% To date, finding the optimal inference method
% % in terms of runtime
% has been
% less important than finding an algorithm easy to modify for \zplus.

MIFD's assessment ranks hypotheses as either
\emph{likely}, \emph{plausible} or \emph{unlikely}. A hypothesis $h$,
conditioned on evidence $\omega$,
is \emph{likely} if
$\kappa(h|\omega) < \kappa(\neg h|\omega)$, \emph{plausible} if $\kappa(h|\omega) = \kappa(\neg h|\omega)$
and \emph{unlikely} otherwise.

The MIFD system requires comparatively few \tkappa parameters given the above
design.  For sensors, we need \fpKappa, and for events (attack and benign)
we need $\kappa(\text{event})$.  In practice, we assign global defaults for
these, based on how generally accurate the input \idses are: for example, we set
these \tkappa\ values so that it takes $n = 2\ \text{or}\ 3$ sensors for us to judge an event
as \emph{likely}, with
the following consistency constraints:
\begin{align}
  \label{eqn:coherence-ineq}
  \kappa(\text{false-positive}),\kappa(\text{benign})
  &< \kappa(\text{attack}) \\
 &< n \cdot \kappa(\text{false-positive}) \nonumber
\end{align}
That is, a single sensor false positive is less surprising than an
  attack event;
a benign event is also less surprising than an attack; and
an attack is less surprising than false positives from $n$
  of the sensors detecting that event.
%
With a few exceptions, this paper assumes $n=3$ sensors, so we
have $0 < \kappa(\text{benign}) < \kappa(\text{attack}) 
    < 3 \kappa(\text{false-positive})$.
% \begin{displaymath}
%   \begin{split}
%     0 &< \kappa(\text{benign-event}) < \kappa(\text{attack-event}) \\
%     &< 3 \kappa(\text{false-positive})
%   \end{split}
% \end{displaymath}
% \noindent
In actual deployments, we then
nudge the false positive rankings up in the rare cases where we have a
particularly good sensor, or down if we have a particularly bad sensor.
We are less likely to adjust attack event probabilities, but
there are exceptions such as scanning (very common), and some complex events.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-12 ids-fusion.tex
Getting under 6,000 words

@@
 often voluminous, stream of reports; fusing them into a coherent 
-overview is left as an exercise for the user. 
+overview is left
+% TODO restore to QR15
+%as an exercise for
+ to
+the user. 
 Ideally, network administrators would have a suite of different \idses active,
-since different \ids approaches --- network (NIDS) vs.\
-host-based (HIDS), signature vs.\ anomaly-based --- bring different
+since different \ids approaches
+% TODO restore to QR15
+% --- network (NIDS) vs.\
+% host-based (HIDS), signature vs.\ anomaly-based --- 
+bring different
 strengths and weaknesses.
@@
 patients.  The probability of having the disease given a positive test result is
-only 1\%, although the probability of the sensor labeling correctly is 99\%.
+only 1\%.
+% %% This next bit seem redundant.
+%, although the probability of the sensor labeling correctly is 99\%.
 {\ids} users regularly either ignore or partially disable them, unable to absorb
-massive streams of false alarms:
-``Users attending an 'ABCs of IDS' event at London's City University yesterday
-  said more the 80 per cent of the alerts they received were false, with one
-  citing 60 alerts he had received about non-existent problems that morning at
-  0300''~\cite{Register}.
+massive streams of false alarms:%
+%% TODO Restore for QR'15
+%``Users attending an 'ABCs of IDS' event at London's City University yesterday
+%  said more the 80 per cent of the alerts they received were false, with one
+%  citing 60 alerts he had received about non-existent problems that morning at
+%  0300''
+~\cite{Register}.
 Recently, Target is reported to have ignored warnings about the
@@
 flow of reports in a small corporate network~\anoncite{goldman:09Scyllarus}.
-Data reduction, in addition to building a comprehensive overview
-and overcoming false positives and negatives, must be a goal of \ids fusion.
-% \begin{figure*}[t]
-%   \begin{tabular}{c}
-%     \ifthenelse{\boolean{blackAndWhite}}{%
-%       \includegraphics[width=.7\textwidth]{figures/funnel-bw.pdf} }{%
-%     \includegraphics[width=.7\textwidth]{figures/funnel.pdf}} \\
-%     \ifthenelse{\boolean{blackAndWhite}}{%
-%         \includegraphics[width=.7\textwidth]{figures/funnel-graph-bw} }{%
-%             \includegraphics[width=.7\textwidth]{figures/funnel-graph} } 
-%   \end{tabular}
-%   \centering
-%   \caption{Scyllarus workload reduction~\anoncite{goldman:09Scyllarus}.}
-%   \label{fig:workloadReduction}
-% \end{figure*}
+Data reduction must join building a comprehensive overview
+and overcoming false positives and negatives as goals of \ids fusion.
 \begin{figure}[t]
@@
   \centering
-  \caption{Scyllarus workload reduction~\anoncite{goldman:09Scyllarus}.}
+  \caption{Scyllarus workload reduction%
+% Squash citation here; cited in sentence referencing figure.
+%~\anoncite{goldman:09Scyllarus}
+.}
   \label{fig:workloadReduction}
@@
 \paragraph{Related work.}
-SecurityFocus has developed the
-Attack Registry and Intelligence Service~\shortcite[ARIS]{ARIS}.
-Its extractor consolidates {\ids} reports from four different
-{\idses} into XML, and presents them in an incident
-console.  However it makes no attempts to fuse the reports, nor weigh
-evidence for and against them.
-MetaSTAT is a fusion system built on a set of STAT-based
-{\idses}~\cite{vigna01:designing}.  
-STAT is signature-based, and detects events by matching
-against extended finite-state event models.  MetaSTAT uses
-finite-state models of across-sensor events to consume events
-from lower-level sensors at a higher level,
-but does not attempt to judge the plausibility of different
+%% TODO Restore for QR'15
+%SecurityFocus has developed the
+%Attack Registry and Intelligence Service~\shortcite[ARIS]{ARIS}.
+%Its extractor consolidates {\ids} reports from four different
+%{\idses} into XML, and presents them in an incident
+%console.  However it makes no attempts to fuse the reports, nor weigh
+%evidence for and against them.
+%
+%% TODO Restore full version for QR'15
+STAT and MetaSTAT~\cite{vigna01:designing}
+use finite-state models to detect and fuse events,
+but do not attempt to judge the plausibility of different
 events.
+%MetaSTAT is a fusion system built on a set of STAT-based
+%{\idses}~\cite{vigna01:designing}.  
+%STAT is signature-based, and detects events by matching
+%against extended finite-state event models.
+%MetaSTAT uses
+%finite-state models of across-sensor events to consume events
+%from lower-level sensors at a higher level,
+%but does not attempt to judge the plausibility of different
+%events.
 EMERALD/eBayes~\cite{raid2001-pac} fusion is the most similar to our systems.
@@
 address the difficult issues of acquiring probability parameters for \ids fusion.
-The Prelude IDS's Correlator allows users to analyze reports sent to Prelude 
-from compatible IDSes~\cite{PreludeCor} following user-provided rules.  Its function is closest to
+The Prelude IDS's Correlator
+%% TODO Restore full version for QR'15
+% allows users to analyze reports sent to Prelude 
+%from compatible IDSes~\cite{PreludeCor} following user-provided
+%rules.  Its function
+is closest to
 our clustering subsystem, but knowledge resides in stateful 
 rules instead of an ontology of attacks. 
-The Arcsight Enterprise Security Manager~\cite{ArcsightEsm},
-also ties correlated IDS reports to an installation's security goals
-and vulnerability information.
+%% TODO Restore full version for QR'15
+%The Arcsight Enterprise Security Manager~\cite{ArcsightEsm},
+%also ties correlated IDS reports to an installation's security goals
+%and vulnerability information.
 

### 2015-03-12 scyllarus-and-mifd.tex
Getting under 6,000 words

@@
 
-Our original system Scyllarus was an open system which performed IDS
+Our original system Scyllarus
+%% TODO Restore for QR'15
+% was an open system which 
+performed IDS
 fusion on a diverse set of third party \idses~\anoncite{goldman:discex2001,goldman:discex2001}.
@@
 MIFD (Model-based Intrusion Fusion and Detection),
-enhances Scyllarus as part of the STRATUS system's
-integration of IDS fusion into a comprehensive suite for cyber defense
+enhances Scyllarus
+%% TODO Restore for QR'15
+% as part of the STRATUS system's integration of
+ to integrate
+IDS fusion into a comprehensive suite for cyber defense
 \anoncite{thayer13:_compar_strat_tactic_respon_cyber_threat}.
@@
 preliminaries, except to mention that they
-involve a substantial amount of data rectification.  There is no widely-accepted
-common data format (the proposed IDMEF standard~\cite{IDMEF} has not caught on), nor
-is there an accepted taxonomy for events or event roles (e.g., the widely used ``source''
-and ``destination'' are not consistently applied).
-These problems are exacerbated by the fact that intrusion detection ``systems''
-are often simply a casually-assembled anthology of rules by different authors running in the same
-framework.
+involve a substantial amount of data rectification
+%% TODO Restore for QR'15
+ to address the lack of standardization for data formats and event taxonomies.
+%.  There is no widely-accepted
+%common data format (the proposed IDMEF standard~\cite{IDMEF} has not caught on), nor
+%is there an accepted taxonomy for events or event roles (e.g., the widely used ``source''
+%and ``destination'' are not consistently applied).
+%These problems are exacerbated by the fact that intrusion detection ``systems''
+%are often simply a casually-assembled anthology of rules by different authors running in the same
+%framework.
 
@@
 may look like an attacker performing reconnaissance.
-What's worse, some of the false positives are correlated: the normal
+Worse, some false positives are correlated: the normal
 activity of a print server may be mis-detected as a scan by several
@@
 introduce benign events, that can compete with attack and (simple) false
-positive explanations for results.\footnote{Where possible, we try to introduce sensors that
-can specifically detect the benign events. To return to our earlier example, we
-might introduce a ``print server sensor'' to help disambiguate detections. We do
-not address this issue in our experiments.}
+positive explanations for results.\footnote{And where possible, we try to introduce sensors that
+can specifically detect the benign events. 
+%% TODO Restore for QR'15
+%To return to our earlier example, we
+%might introduce a ``print server sensor'' to help disambiguate detections. We do
+%not address this issue in our experiments.
+}
 % FIXME: Add a figure 2(c) -- or separately numbered figure -- where we have a
@@
 
-MIFD also has \emph{complex} events that have component sub-events, such as a
-phishing attack with component events phishing email and drive-by-download.
-Our current experiments do not address the role of complex events, so we will
-not discuss them further.
+%% TODO Restore for QR'15 --- we'd earlier discussed
+%% removing this paragraph anyway.
+%MIFD also has \emph{complex} events that have component sub-events, such as a
+%phishing attack with component events phishing email and drive-by-download.
+%Our current experiments do not address the role of complex events, so we will
+%not discuss them further.
 

### 2015-03-12 zplus.tex
Getting under 6,000 words

@@
 calculations.  He recommends these patterns be used even when precise numbers are
-not available, and that is what MIFD does.  As outlined above, MIFD treats
+not available, which MIFD does.
+%% TODO Restore for QR'15
+%As outlined above,
+MIFD treats
 \ids fusion as an abductive problem, formalized using Bayes nets.
-But precise values for the parameters of these Bayes nets ---
-probabilities of attacks, false positives, false negatives, etc. ---
+But precise values for the parameters of these Bayes nets
+%% TODO Restore for QR'15
+% --- probabilities of attacks, false positives, false negatives, etc. ---
 are \emph{never} available to us: populations of
@@
 Goldszmidt and
-Pearl's \zplus{}~\shortcite{Goldszmidt:96},
+Pearl's \zplus{}~\shortcite{Goldszmidt:96}
 shares the
-basic structure of normal probability theory, but abstracts the actual
+basic structure of normal probability theory but abstracts the actual
 probabilities.
@@
 of the polynomial's leading term.
-As we mentioned earlier, \zplus is similar to the ``big-$O$'' method used
+%% TODO Restore for QR'15
+%As we mentioned earlier, 
+\zplus is similar to the ``big-$O$'' method used
 in analysis of algorithms.
-This semantics gives \zplus{} a
-qualitative flavor by providing a ``ladder'' of events of qualitatively different
+With this semantics \zplus{}
+provides a ``ladder'' of events of qualitatively different
 orders of likelihood.
-Note that \zplus is very different from Wellman's qualitative probability
-networks~\cite{Wellman:90}, whose abstraction is based on direction of change, rather than order
-of magnitude relations ``at rest,'' as it were.
-Wellman's approach is more relevant when choosing interventions (as in his work
-on medical diagnosis and treatment), and less for fusion.
+%% TODO Restore for QR'15
+%Note that \zplus is very different from Wellman's qualitative probability
+%networks~\cite{Wellman:90}, whose abstraction is based on direction of change, rather than order
+%of magnitude relations ``at rest,'' as it were.
+%Wellman's approach is more relevant when choosing interventions (as in his work
+%on medical diagnosis and treatment), and less for fusion.
 
@@
 Under \zplus\ we may apply the normal operation
-of probability theory --- conditionalization, Bayes' law, etc. --- but
-the arithmetic operations we use for them change.  Rather than multiplying
-probabilities, we add degrees of surprise; rather than adding
-probabilities, we take the minimum surprise.  Goldszmidt and
+of probability theory
+%% TODO Restore for QR'15
+% --- conditionalization, Bayes' law, etc. ---
+but the arithmetic operations we use for them change.  Rather than multiplying
+probabilities, we add degrees of surprise%
+%% TODO Restore for QR'15, but it's in the table anyway
+%; rather than adding
+%probabilities, we take the minimum surprise%
+.  Goldszmidt and
 Pearl~\shortcite{Goldszmidt:96} provide the following substitutions
@@
 handle common special cases.
-To date (somewhat to our disappointment!) optimizing the general \zplus
-inference method has not been key to system performance.
+%% TODO Restore for QR'15
+%To date (somewhat to our disappointment!) optimizing the general \zplus
+%inference method has not been key to system performance.
+%
 % To date, finding the optimal inference method
@@
 particularly good sensor, or down if we have a particularly bad sensor.
-We are less likely to adjust attack event probabilities, but
-there are exceptions such as scanning (very common), and some complex events.
+%% TODO Restore for QR'15
+%We are less likely to adjust attack event probabilities, but
+%there are exceptions such as scanning (very common), and some complex events.
 

### 2015-03-12 intro.tex
Getting under 6,000 words
Spurious comma

@@
 In previous work we developed a technique for IDS
-fusion, deployed in the Scyllarus system~\anoncite{goldman:09Scyllarus}, and its
+fusion, deployed in the Scyllarus system~\anoncite{goldman:09Scyllarus} and its
 successor MIFD~\anoncite{stratus:saso12}.
@@
 problem, as implemented in the Scyllarus and MIFD systems.  We describe our
-experimental designs, present the results and conclude with some proposals for
+experimental designs, present the results, and conclude with some proposals for
 future work.

### 2015-03-12 design.tex
Minor edits to the text.
One note: as I read this, I realized that the reader doesn't need to
know that we have event and sensor PROTOTYPES instead of TYPES. That's
an implementation detail that comes from an oddity of PRAXIS. So I
suppressed mention of "prototypes."

@@
 For each experiment we create a \emph{configuration},
-parameters specifying how a set of events and sensors (the
-\emph{setting}) is generated.  A
+a set of parameters specifying how a set of events and sensors, a
+\emph{setting}, is generated.  A
 setting comprises a set of possible attacks, a set
@@
 % below.
-In each experiment we generated 1,000 settings with
-100 runs per setting.
+In each experiment, for each configuration, we generated 1,000 settings
+and conducted 100 runs per setting.
 %
@@
 \begin{compactitem}
-\item \numEventProtosDetected\ --- The total number of events (both
+\item \numEventProtosDetected\ --- The total number of event types (both
   benign and attacks) to be
@@
 When creating a setting, we first consult
-the \numAttacks\ parameter to determine the number of attack prototypes.
-Each prototype is associated with an
-occurrence probability given by (or sampled for each prototype from)
-\attackKappa. The number of sensors is not fixed, but instead depends
+the \numAttacks\ parameter to determine the number of attack types.
+Each type is associated with an
+occurrence probability dictated by the 
+\attackKappa parameter. The number of sensors is not fixed, but instead depends
 on several configuration parameters. For each attack, we keep track of
@@
 events. Each sensor is assigned at least one attack, plus
-additional attacks up to its total number of events as consecutive
-samples of a Bernoulli random variable with success probability
-\sensorOverlap\ return 1.  After all sensor prototypes are
+additional attacks up to its total number of events based on a 
+Bernoulli random variable with parameter
+\sensorOverlap.  After all sensor types are
 created and assigned attacks, we assign benign events to the sensors
@@
 %% \item If we are given \numBenignEvents, then we create that total
-%%   number of benign event prototypes, and distribute them among the
+%%   number of benign event types, and distribute them among the
 %%   sensors using a \sensorToBenignRatio\ which allows each sensor's
@@
 %%   If we are given a \sensorToBenignRatio, then we create the
-%%   necessary number of benign event prototypes to satisfy both
+%%   necessary number of benign event types to satisfy both
 %%   \sensorToBenignRatio\ and each sensor's total event count.
 %% \end{itemize}
-We create enough benign event prototypes to satisfy
+We create enough benign event types to satisfy
 both \sensorToBenignRatio\ and each sensor's total event count, and
@@
 % events which occur at each location by considering each attack and
-% benign prototype. For each prototype, if a sample of a Bernoulli
+% benign type. For each type, if a sample of a Bernoulli
 % random variable whose success probability is the occurrence
-% probability of the prototype is 1, then the ground truth aspect will
-% contain a sensor report with that prototype and location.
+% probability of the type is 1, then the ground truth aspect will
+% contain a sensor report with that type and location.
 % %
 % For sensor reports at each location, we consider each sensor report
-% prototype of the setting. If a prototype indicates that the sensor can
+% type of the setting. If a type indicates that the sensor can
 % detect one of the events which we have generated for that location,
@@
 %   Bernoulli random variable whose success probability is the false
-%   negative probability of the report prototype. If \emph{any} of these
+%   negative probability of the report type. If \emph{any} of these
 %   samples are 0 --- that is, if for any event we do \emph{not} have a
 %   false negative --- then the ground truth aspect will contain a
-%   sensor report with that prototype and location.
+%   sensor report with that type and location.
 % \item If there are no events at a location which could trigger a
-%   report of that prototype, then we consult a Bernoulli random
+%   report of that type, then we consult a Bernoulli random
 %   variable whose success probability is the false positive probability
-%   of the sensor report prototype; if it returns 1, then the ground
-%   truth aspect will contain a sensor report with that prototype and
+%   of the sensor report type; if it returns 1, then the ground
+%   truth aspect will contain a sensor report with that type and
 %   location.

### 2015-03-12 intro.tex
Minor edits to the text.
One note: as I read this, I realized that the reader doesn't need to
know that we have event and sensor PROTOTYPES instead of TYPES. That's
an implementation detail that comes from an oddity of PRAXIS. So I
suppressed mention of "prototypes."

@@
 probability scheme when used for diagnosis and information fusion.
-In cyber defense, an ``intrusion detection system (IDS) is a device or
+In cyber defense, an ``\ldots intrusion detection system (IDS) is a device or
 software application that monitors network or system activities for malicious
@@
 We also
-drastically reduce false alarm rates, the bane of intrusion
+dramatically reduce false alarm rates, the bane of intrusion
 detection systems, reducing the flood of incoming reports by multiple
 orders of magnitude.
-Unfortunately, it is difficult to draw crisp, \emph{general} conclusions based on
+Unfortunately, we cannot draw crisp, \emph{general} conclusions based on
 field evaluations alone.
@@
 work~\cite{google-scholar-query}, none of the few which detail
-applications of \zplus\ reasoning conduct such investigations.
+applications of \zplus\ reasoning (as opposed to theoretical investigations of
+the logic) conduct such investigations.
 %

### 2015-03-12 scyllarus-and-mifd.tex
Minor edits to the text.
One note: as I read this, I realized that the reader doesn't need to
know that we have event and sensor PROTOTYPES instead of TYPES. That's
an implementation detail that comes from an oddity of PRAXIS. So I
suppressed mention of "prototypes."

@@
 fusion on a diverse set of third party \idses~\anoncite{goldman:discex2001,goldman:discex2001}.
-Its successor
+Its successor,
 MIFD (Model-based Intrusion Fusion and Detection),
@@
 %% TODO Restore for QR'15
- to address the lack of standardization for data formats and event taxonomies.
+ to address the lack of standardization in data formats and event taxonomies.
 %.  There is no widely-accepted
@@
 
-Core fusion inference begins with
-\textbf{Clustering:} From the
+{\ids} fusion proper begins with
+\textbf{Clustering.} From the
 sensor reports, MIFD generates \emph{event hypotheses} to represent the
@@
 introduce benign events, that can compete with attack and (simple) false
-positive explanations for results.\footnote{And where possible, we try to introduce sensors that
+positive explanations for results.\footnote{Where possible, we try to introduce sensors that
 can specifically detect the benign events. 
@@
 
-The final inference step is \textbf{Assessment:} MIFD assigns a degree of belief
+The second inference step is \textbf{Assessment,} in which MIFD assigns a degree of belief
 to each of the event hypotheses.
@@
 %%
-Scyllarus and MIFD are intended as long-running online services, not batch processes.
+
+Note that Scyllarus and MIFD are intended as long-running online services, not batch processes.
 Scyllarus has run for months at a time.
@@
 cycle to perform assessment.
+Scyllarus also periodically flushes older events and reports out of working
+memory (they persist in a database).
+
 %% FIXED Although this next sentence is a true fact, I'm not sure it

### 2015-03-12 zplus.tex
No longer need to divide anon. from other citations.
Minor edits to the text.
One note: as I read this, I realized that the reader doesn't need to
know that we have event and sensor PROTOTYPES instead of TYPES. That's
an implementation detail that comes from an oddity of PRAXIS. So I
suppressed mention of "prototypes."

@@
 calculations.  He recommends these patterns be used even when precise numbers are
-not available, which MIFD does.
+not available.
 %% TODO Restore for QR'15
@@
 non-stationary, etc.  Moreover these difficulties are not solvable via machine learning:
-Sommer and Paxson~\shortcite{Sommer10outsidethe} discuss the challenges \ids pose to ML.
+Sommer and Paxson~\shortcite{Sommer10outsidethe} discuss the challenges \idses pose to ML.
 
@@
 %% breaching anonymity. [2014/08/24:rpg]
-ATMS~\cite{Forbus+DeKleer:1993}; see \cite{Charniak:88,Poole:93,provan-ijcai89}
-and \anoncite{goldman:09Scyllarus} for this encoding.
+ATMS~\cite{Forbus+DeKleer:1993}; see
+%% For anonymous:
+%\cite{Charniak:88,Poole:93,provan-ijcai89}
+%and \anoncite{goldman:09Scyllarus}
+%% De-anonymized:
+\cite{Charniak:88,Poole:93,provan-ijcai89,goldman:09Scyllarus}
+%
+for this encoding.
 This implementation is not especially efficient, but we have found

### 2015-03-20 conclusions.tex
Echo absence of quantitative model in conclusions per Scott's suggestion

\section{Conclusions}
\label{sec:conclusions}

The results presented in this paper clarify why Scyllarus and
MIFD have been successful in practical deployments.
From their earliest design, the qualitative probabilities of \zplus\ 
seemed like a good choice since we will never have precise numeric
probabilities for sensor behaviors and event likelihoods, but
on the other hand it is quite reasonable and natural to characterize
them qualitatively.
Here we have shown that our use of \zplus\ is not sensitive to
the actual probabilities of events, and that its performance degrades
gracefully.
The results also justify the claim that by combining the output of multiple sensors, even
very noisy sensors, \ids fusion can tame the high false
positive rates that plague the field of intrusion detection.
Since our results predict the circumstances under which IDS fusion will work and
will fail, they can also be used to inform the design and deployment of IDSes
for effective incorporation in a fusion system.
Finally, our results should encourage prospective users
of qualitative schemes based on probabilistic reasoning,
and promote deeper examination of % the usefulness and limits of
such systems.
\hide{In current work, we are extending our evaluation to consider MIFD's behavior
when topology is critical.  Cyber attacks
spread through network links, both actual
communications links and superimposed networks induced by protocols.
}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2015-03-25 conclusions.tex
Revert "Echo absence of quantitative model in conclusions per Scott's suggestion"
This reverts commit 90c60f075818473700058622d8ccae28f7c07fbe.

The patch did address Scott's point, but in a way that introduced a five line
run-on sentence.  Will try to tighten this up.

@@
 MIFD have been successful in practical deployments.
-From their earliest design, the qualitative probabilities of \zplus\ 
-seemed like a good choice since we will never have precise numeric
-probabilities for sensor behaviors and event likelihoods, but
-on the other hand it is quite reasonable and natural to characterize
-them qualitatively.
-Here we have shown that our use of \zplus\ is not sensitive to
+They show that the qualitative scheme they use is not sensitive to
 the actual probabilities of events, and that its performance degrades

### 2015-03-26 conclusions.tex
Substantially expanded conclusions.
Tried to address all of Scott's suggestions in depth.

@@
 gracefully.
-The results also justify the claim that by combining the output of multiple sensors, even
-very noisy sensors, \ids fusion can tame the high false
-positive rates that plague the field of intrusion detection.
+
+We would like to remind the reader that imprecise probabilities 
+are forced upon us by the nature of the cyber intrusion detection domain: it's
+not just a matter of not having the right machine learning technique.  The
+true probabilities are non-stationary, they involve an adversarial process,
+vary from location to location, and are difficult if not impossible to learn
+because of the absence of valid labeled training data.
+The use of a small number of qualitatively distinct likelihood levels also aids
+us in the knowledge engineering process.
+
+In practice, we also find that the results of \zplus calculations are easy to
+understand.  We have found this through experience explaining the output of our
+systems to users, and through experience debugging.  When we draw out the Bayes
+networks, we can think of using assumptions to assemble the ``cheapest''
+explanation for a set of observations.  This can be simpler to understand than
+the exact computations in a Bayes net.
+
+Our experimental results also justify the claim that by combining the output of
+multiple sensors, even very noisy sensors, \ids fusion can tame the high false
+positive rates that plague the field of intrusion detection.  These results are
+somewhat independent from the question of the adequacy of the qualitative
+calculus.  In a system that performed diagnosis/fusion using conventional
+probability theory, it can be seen that multiple sensors that fail independently
+will tend to perform well:  it's easier to drive the probability of error down
+by multiplying failure probabilities ($p^n$) than to try to drive down a single
+sensor's $p$ of failure.  However, our results show that substituting \zplus for
+conventional probability theory preserves this desirable characteristic of
+probabilistic reasoning. Our results also show resistance to a moderate degree
+of correlation in sensor failures when using \zplus.
+
 Since our results predict the circumstances under which IDS fusion will work and

### 2015-06-26 ids-fusion.tex
The Prelude citation, maybe

@@
 of a 99\% accurate test detecting a disease that strikes only 1/10,000
-patients.  The probability of having the disease given a positive test result is
+patients.  
+Although the sensor is accurate, 
+the probability of having the disease given a positive test result is
 only 1\%.
@@
 %, although the probability of the sensor labeling correctly is 99\%.
-{\ids} users regularly either ignore or partially disable them, unable to absorb
-massive streams of false alarms:%
+
+Users regularly ignore or disable their {\idses}, unable to absorb
+massive streams of false alarms.
 %% TODO Restore for QR'15
@@
 %  0300''
-~\cite{Register}.
+% ~\cite{Register}.
 Recently, Target is reported to have ignored warnings about the
 data breach that resulted in theft of millions of credit
-cards~\cite{TargetIgnoredIDS}:
+card records~\cite{TargetIgnoredIDS}:
 ``They are bombarded with alerts. They get so many that they just don't respond
@@
 \ref{fig:workloadReduction} gives a sense of the gravity of this
-problem, showing how our earlier system was able to winnow
+problem. It shows how our earlier system was able to winnow
 the 
 flow of reports in a small corporate network~\anoncite{goldman:09Scyllarus}.
-Data reduction must join building a comprehensive overview
-and overcoming false positives and negatives as goals of \ids fusion.
+% Data reduction must join building a comprehensive overview
+% and overcoming false positives and negatives as goals of \ids fusion.
 \begin{figure}[t]
@@
 % allows users to analyze reports sent to Prelude 
-%from compatible IDSes~\cite{PreludeCor} following user-provided
+%from compatible IDSes~\cite{PreludeCorNew} following user-provided
 %rules.  Its function
 is closest to
-our clustering subsystem, but knowledge resides in stateful 
-rules instead of an ontology of attacks. 
+our clustering subsystem, but its knowledge resides in stateful 
+rules instead of an ontology of attacks~\cite{PreludeCorNew}. 
 %% TODO Restore full version for QR'15

### 2015-06-26 design.tex
Learned or hand-mode domains? --- over to RPG

@@
 
-In all of these experiments, we define sensor arrangements, and probabilities for attacks, confounding
-benign events, and sensor false positives and false negatives. We sample from
+In all of these experiments, we define sensor arrangements as well as the probabilities for attacks, confounding
+benign events, sensor false positives and false negatives. We sample from
 them to find ground truth and sensor report sets, run MIFD on the sensor
@@
 a set of parameters specifying how a set of events and sensors, a
-\emph{setting}, is generated.  A
+\emph{setting}, is generated by sampling random variables and
+comparing to these parameters.  A
 setting comprises a set of possible attacks, a set

### 2015-06-26 ids-fusion.tex
Prelude citation really fixed.
This paragraph was a little convoluted, I think as a consequence of anonymous citation.

@@
 \end{compactenum}
-In this paper we will focus on the second of these two questions, assuming a
-solution to the first.  For discussions of the first question, see
-\anoncite{goldman:09Scyllarus}.
+Goldman and Harp discuss the first of these two
+questions~\shortcite{goldman:09Scyllarus}.  In this paper we will
+focus on the second question, assuming a solution to the first.
 
@@
 % allows users to analyze reports sent to Prelude 
-%from compatible IDSes~\cite{PreludeCorNew} following user-provided
+%from compatible IDSes~\cite{PreludeCor} following user-provided
 %rules.  Its function
@@
 our clustering subsystem, but its knowledge resides in stateful 
-rules instead of an ontology of attacks~\cite{PreludeCorNew}. 
+rules instead of an ontology of attacks~\cite{PreludeCor}. 
 %% TODO Restore full version for QR'15

### 2015-06-26 ids-fusion.tex
Flavors of IDS --- kicked in a sentence to clarify

@@
 false positives and false negatives.
+This tradeoff applies for all of the different varieties of IDS,
+for example NIDS or HIDS, signature-based, anomaly-detecting, various
+hybrids.  We do not discuss the varieties of \idses here;
+we are not concerned with the mechanics of clustering reports, but
+rather with the subsequent assessment of hypothesized events.
 %%

### 2015-06-26 scyllarus-and-mifd.tex
Reference to ontology of attacks

@@
 \textbf{Clustering.} From the
-sensor reports, MIFD generates \emph{event hypotheses} to represent the
+sensor reports,
+and based on an ontology detailing how (and with what probability)
+sensor reports may correspond to events,
+MIFD generates \emph{event hypotheses} to represent the
 underlying events of interest that may have caused the sensor reports.  A

### 2015-07-14 ids-fusion.tex
Indicated that the funnel is for one day.
Mods relevant to todo Review 3: Flavors of IDS

--- 2015-06-26 ids-fusion.tex
@@
 bring different
-strengths and weaknesses.
+strengths and weaknesses, and have different ``fields of view.'' For example, local
+privilege escalation attacks may be invisible to a NIDS; HIDS systems are more
+accurate for many purposes, but impose higher administrative burdens because
+they must be scattered around the defended network, rather that centrally
+deployed like NIDS.  Signature-based IDSes provide fewer false positives, but
+are more likely to be blind to zero-day attacks than anomaly-based systems.
 %%
@@
 %%
-Briefly, more accurate \idses have significantly higher software
-administration burdens, and there is an inevitable tradeoff between
-false positives and false negatives.
+% Briefly, more accurate \idses have significantly higher software
+% administration burdens, and there is an inevitable tradeoff between
+% false positives and false negatives.
+% This tradeoff applies for all of the different varieties of IDS,
+% for example NIDS or HIDS, signature-based, anomaly-detecting, various
+% hybrids.
+
+% clarification relevant to [[file:todo.txt::*Review%203:%20Flavors%20of%20IDS][Review 3: Flavors of IDS]]
+We will not discuss the varieties of \idses further here -- there is a
+voluminous literature on them.   The \emph{precise} nature of \ids weaknesses and
+strengths is not critical to our investigation here, but the fact that there
+\emph{are} such strengths and weaknesses is what motivates our information
+fusion, and difficulties in modeling the events and sensors is what motivates
+our use of qualitative methods.
 %%
@@
 the 
-flow of reports in a small corporate network~\anoncite{goldman:09Scyllarus}.
+flow of reports in a small corporate network over a day of operation~\anoncite{goldman:09Scyllarus}.
 % Data reduction must join building a comprehensive overview

### 2015-07-14 ids-fusion.tex

@@
 %, although the probability of the sensor labeling correctly is 99\%.
+
+% New paragraph to address
+% [[file:todo.txt::*Review%201:%20learned%20or%20hand-made%20domains?][Review 1: learned or hand-made domains?]]
+% and [[file:todo.txt::*Review%201:%20Hand-made%20vs.%20generated][Review 1: Hand-made vs. generated]]
+Some of our reviewers expressed the concern that we would have difficulty
+\emph{modeling} the rare events in \ids fusion.  This problem does not, in fact,
+arise, for three reasons.  One is that these attack events are not rare in the
+world: they are rare with respect to our sampling rate.  As we point out above,
+if we are sampling individual packets, the probability of any attack is
+vanishingly small, but that doesn't mean that attacks are rare at the
+``macroscopic'' level.  Second, our causal models are limited to models of
+sensing and fusion, part/whole, and class/subclass relations.  These models are
+very simply generated by rules.  Finally, the \ids developers provide us initial
+ontologies of events implicitly with their rules.  Our knowledge engineering
+aims at smoothing inconsistencies between these implicit ontologies, and
+extending causal models by identifying cases where \ids developers have
+provided incorrect intuitions about the events their systems detect.
+At any rate, this concern is not relevant to the evaluation of the uncertainty
+calculus, which is our focus in this paper.
 

### 2015-07-14 zplus.tex
Limitiation to three classes --- text in
Typo in table
Address question about only three levels of posterior surprise.

@@
   P(\psi) = \sum_{e \in \psi} P(e) & \kappa(\psi) =
-  \operatorname{min}_{e \in \psi} \kappa(\phi) \\ \hline
+  \operatorname{min}_{e \in \psi} \kappa(e) \\ \hline
   P(\psi) + P(\neg \psi) = 1 & \kappa(\psi) = 0 \vee \kappa(\neg \psi) = 0
@@
 
-MIFD's assessment ranks hypotheses as either
+% Extended this to address [[file:todo.txt::*Review%203:%20Limitation%20to%20three%20classes][Review 3: Limitation to three classes]]
+\mifd's assessment ranks hypotheses as either
 \emph{likely}, \emph{plausible} or \emph{unlikely}. A hypothesis $h$,
@@
 and \emph{unlikely} otherwise.
+That is, a hypothesis is likely (unlikely) if some maximally likely scenario labels it as
+true (false) , and no maximally likely scenario labels it as false (true).  A
+hypothesis is plausible, in any other situation.
+Note that this restriction to three classes applies to \mifd's output,
+not its inputs --- the \tkappa\ values of \mifd's underlying model
+range over arbitrary natural numbers as needed to distinguish
+qualitatively distinct likelihoods.
+It would be possible to alter the assessment process to provide more classes of
+likelihood, using the difference in kappa ranks between the maximally likely
+scenario in which a hypothesis is labeled as true (false) and the one in which
+it is labeled as false (true).  We have not seen the need to do so and, indeed,
+worry that additional precision would come at the cost of validity.
 

### 2015-07-14 zplus.tex
Squash some commas.

@@
 That is, a hypothesis is likely (unlikely) if some maximally likely scenario labels it as
-true (false) , and no maximally likely scenario labels it as false (true).  A
-hypothesis is plausible, in any other situation.
+true (false), and no maximally likely scenario labels it as false (true).  A
+hypothesis is plausible in any other situation.
 Note that this restriction to three classes applies to \mifd's output,
@@
 scenario in which a hypothesis is labeled as true (false) and the one in which
-it is labeled as false (true).  We have not seen the need to do so and, indeed,
+it is labeled as false (true).  We have not seen the need to do so, and indeed
 worry that additional precision would come at the cost of validity.

### 2015-07-14 ids-fusion.tex
Editing --- simpler sentence.
Not sure if there's an actual change here --- git funniness.

@@
 overview is left
-% TODO restore to QR15
+% WONTFIX restore to QR15
 %as an exercise for
@@
 since different \ids approaches
-% TODO restore to QR15
+% WONTFIX restore to QR15
 % --- network (NIDS) vs.\
@@
 Some of our reviewers expressed the concern that we would have difficulty
-\emph{modeling} the rare events in \ids fusion.  This problem does not, in fact,
-arise, for three reasons.  One is that these attack events are not rare in the
+\emph{modeling} the rare events in \ids fusion.  There are three
+reasons when this problem does not
+arise.  One is that these attack events are not rare in the
 world: they are rare with respect to our sampling rate.  As we point out above,
@@
 massive streams of false alarms.
-%% TODO Restore for QR'15
-%``Users attending an 'ABCs of IDS' event at London's City University yesterday
-%  said more the 80 per cent of the alerts they received were false, with one
-%  citing 60 alerts he had received about non-existent problems that morning at
-%  0300''
-% ~\cite{Register}.
+``Users attending an 'ABCs of IDS' event at London's City University yesterday
+  said more the 80 per cent of the alerts they received were false, with one
+  citing 60 alerts he had received about non-existent problems that morning at
+  0300''~\cite{Register}.
 Recently, Target is reported to have ignored warnings about the

### 2015-07-15 scyllarus-and-mifd.tex
Remove completed FIXME.
Clean a \cite that had the same reference repeated twice
i.e. \cite{xyz,xyz} ===> \cite{xyz}

Also squash some commas.

@@
 performed IDS
-fusion on a diverse set of third party \idses~\anoncite{goldman:discex2001,goldman:discex2001}.
-Its successor,
-MIFD (Model-based Intrusion Fusion and Detection),
+fusion on a diverse set of third party \idses~\cite{goldman:discex2001}.
+Its successor
+MIFD (Model-based Intrusion Fusion and Detection)
 enhances Scyllarus
@@
 might be caused by two different underlying events.
+Finally, Figure~\ref{fig:bnets}(c) shows a more complex example where an event
+has two component sub-events.
 %% DONE: cut the following sentence because its meaning wasn't
@@
 }
-% FIXME: Add a figure 2(c) -- or separately numbered figure -- where we have a
-% benign event, explaining two of the sensor reports, as well as the two attack
-% events. 
+
 Another problem with indirect sensing is that the sensors are often imprecise:
@@
   \includegraphics[width=.267\columnwidth]{figures/bnets2-cropped.pdf} 
-  \\ (a) & (b)
+  \\ (a) & (b) \\
+  \end{tabular}
+  \newline{}
+  \begin{tabular}{c}
+  \includegraphics[width=.4\columnwidth]{figures/bnets3-cropped.pdf} 
+    \\ (c)
   \end{tabular}
@@
   \label{fig:bnets}
+
 \end{figure}

### 2015-07-15 ids-fusion.tex
Quotation mark facing the wrong way.

@@
 massive streams of false alarms.
-``Users attending an 'ABCs of IDS' event at London's City University yesterday
+``Users attending an `ABCs of IDS' event at London's City University yesterday
   said more the 80 per cent of the alerts they received were false, with one
